---
title: "Agentic Academic Researcher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Academic Researcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates a multi-stage academic research pipeline using `HydraR`. We define a linear workflow designed to synthesize academic literature into a final report using the **Gemini API**.

The pipeline comprises three agents:
1.  **Searcher**: Identifies relevant papers for a given topic.
2.  **Summarizer**: Extracts key findings from the papers.
3.  **Compiler**: Formats the findings into a professional literature review.

## Declarative Workflow Definition

We use a YAML file to define the graph structure, the roles of each agent, and the prompt templates. This approach keeps the R code clean and focused on execution.

```{r load}
library(HydraR)

# Load the workflow definition from YAML
wf <- load_workflow("academic_research.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Pipeline

The `plot()` method allows us to verify the architecture.

```{r plot}
dag$plot()
```

## Execution

We run the DAG with a research topic. In this example, we use the default topic from the YAML: `CRISPR Gene Editing`.

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY in the environment
final <- dag$run(wf$initial_state)

# Display the final literature review
cat(final$state$get("Compiler"))
```

## Resilience and Checkpointing

For longer research tasks, you can attach a `Checkpointer` (e.g., `DuckDBSaver`) to the `run()` method. This ensures that if the process is interrupted, you can resume from the last successful node without re-running the entire pipeline.

---
<!-- APAF Bioinformatics | academic_research.Rmd | Approved | 2026-04-03 -->
