---
title: "Advanced Network Logic Circuitry"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Network Logic Circuitry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

`HydraR` supports complex "Network Logic Circuitry" that goes beyond simple Directed Acyclic Graphs. This includes dynamic routing, self-healing networks via error edges, and parallel isolation.

## Core Concepts

1.  **Dynamic Routing**: Use a `Router` node to decide which path to follow at runtime based on the current state.
2.  **Error Handling**: Define `error_edges` to automatically switch to recovery nodes if a task fails.
3.  **Declarative Control**: Define these complex relationships in YAML, keeping your R code strictly focused on high-level orchestration.

## Loading the Circuit

We load a workflow that demonstrates a "High Risk" vs "Low Risk" path based on a `mode` parameter.

```{r setup}
library(HydraR)

# Load the advanced circuitry definition
wf <- load_workflow("advanced_circuitry.yml")

# Spawn the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Switch and Fallback

Notice the **Mode Switch** router and the fallback from **High Risk Path** to **Recovery Node** (this would be represented by an error edge).

```{r plot}
dag$plot()
```

## Running the Scenario

In this scenario, we set `simulate_fail = TRUE` to witness the self-healing capability.

```{r run}
# Run the DAG with 'fast' mode and 'simulate_fail = TRUE'
final <- dag$run(wf$initial_state)

# Check the results
print(final$state$get("finalize_run"))
```

## Resilience Features

-   **Router Nodes**: Allow for branching that is decided by logic, not just hardcoded paths.
-   **Error Edges**: Ensure that critical failures don't crash the entire workflow but instead trigger cleanup or recovery routines.
-   **State Isolation**: Combined with `use_worktrees = TRUE`, these branches can run in complete isolation from one another.

---
<!-- APAF Bioinformatics | advanced_circuitry.Rmd | Approved | 2026-04-03 -->
