---
title: "Agentic Blog Writer: Loops and Feedback"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Blog Writer: Loops and Feedback}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates the **cyclic feedback** pattern in `HydraR`. Content creation often involves iterative refinement between a writer and an editor. We model this as a DAG with a mix of linear execution and conditional loops using the **Gemini API**.

## The Workflow Architecture

1.  **Outliner**: Generates a structured outline based on a topic.
2.  **Drafter**: Writes the draft based on the outline and incorporates any feedback from the Editor.
3.  **Editor**: Reviews the draft for SEO and quality. If it fails to meet standards, it triggers a loop back to the Drafter.

## Declarative Configuration

We define the looping logic and agent roles in a single YAML file.

```{r load}
library(HydraR)

# Load the workflow from YAML
wf <- load_workflow("blog_writer.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Feedback Loop

Notice the back-edge from **Editor** to **Drafter**.

```{r plot}
dag$plot()
```

## Running the Iterative Scenario

The `run()` method will continue executing until either the "Approved" condition is met or the `max_steps` limit is reached.

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY
final <- dag$run(wf$initial_state, max_steps = 10)

# Display the final draft and editor's approval
cat("Final Output:\n", final$state$get("Drafter"), "\n")
cat("Editor Decision:", final$state$get("Editor"), "\n")
```

## Resilience

If the loop runs too long or the API connection is severed, `HydraR` can pause execution. By attaching a checkpointer, you can resume the writing process from the exactly where it left off, preserving all previous drafts and feedback history.

---
<!-- APAF Bioinformatics | blog_writer.Rmd | Approved | 2026-04-03 -->
