---
title: "Agentic Personalized Shopping Assistant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Personalized Shopping Assistant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates how to model an interactive, multi-agent feedback loop using `HydraR`. We simulate a **Personalized Shopping Assistant** where a recommendation engine interacts with a mock user.

## The Interaction Model

The workflow consists of two primary roles:
1.  **Store Concierge**: Acts as the intelligent proxy that searches the catalogue based on preferences.
2.  **Customer Proxy**: Represents the user feedback loop. It critiques the recommended item, providing refinement feedback or final approval.

## Declarative Feedback Loop

We define the looping logic and agent roles in a YAML file.

```{r load}
library(HydraR)

# Load the shopping assistant workflow
wf <- load_workflow("customer_agent.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Interaction

The graph shows the recursive loop between the **Store Concierge** and the **Customer Proxy**.

```{r plot}
dag$plot()
```

## Running the Scenario

The DAG will continue to iterate, with the `Shopper` refining its recommendation based on the `UserProxy`'s feedback, until the customer is satisfied (or the `max_steps` limit is reached).

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY
final <- dag$run(wf$initial_state, max_steps = 10)

# Display the final shopping journey result
cat("Final Recommendation:", final$state$get("Shopper"), "\n")
cat("User Feedback:", final$state$get("UserProxy"), "\n")
```

## Summary

The `Customer Agent` pattern is a powerful way to use LLMs to simulate real-world user interactions and refine outputs through automated, intelligent critiquing. Using `HydraR`'s declarative framework, these complex loops are easy to define and audit.

---
<!-- APAF Bioinformatics | customer_agent.Rmd | Approved | 2026-04-03 -->
