---
title: "Agentic Data Science: AutoML Optimization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Data Science: AutoML Optimization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

In machine learning workflows, hyperparameter tuning is inherently iterative. We can map this paradigm to an `AgentDAG` using the **Gemini API**. This vignette demonstrates how to define an AutoML optimization loop that includes data preprocessing, LLM-driven parameter recommendations, and automated threshold evaluation.

## The Optimization Pipeline

The workflow contains an initial data setup phase followed by a cyclic model optimization phase:
1.  **Data Preprocessor**: Ingests and prepares the raw data for analysis.
2.  **AutoML Optimizer**: Uses an LLM to recommend hyperparameters based on previous performance.
3.  **Threshold Guard**: Assesses the model's accuracy. If it falls below the target threshold, it loops back to the Optimizer with refinement feedback.

## Declarative Workflow

We define the entire optimization loop in a YAML file.

```{r load}
library(HydraR)

# Load the AutoML workflow from YAML
wf <- load_workflow("data_science.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Optimization Loop

The chart clearly shows the feedback loop from the **Threshold Guard** back to the **AutoML Optimizer**.

```{r plot}
dag$plot()
```

## Running the Scenario

The DAG intelligently executes the data prep before dropping into an iterative optimization loop until the target metric (85% accuracy) is satisfied!

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY
final <- dag$run(wf$initial_state, max_steps = 15)

# Inspect the final tuning result
cat("Final Configuration:", final$state$get("ModelTrainer"), "\n")
cat("Final Accuracy Status:", final$state$get("eval_message"), "\n")
```

## Summary

By combining deterministic R logic (preprocessing, evaluation) with LLM agents (optimization), `HydraR` provides a flexible and powerful framework for building complex, self-correcting data science pipelines.

---
<!-- APAF Bioinformatics | data_science.Rmd | Approved | 2026-04-03 -->
