---
title: "Distributed Consensus: Secure Messaging API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Distributed Consensus: Secure Messaging API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates how to use the `HydraR` messaging API to simulate a **Distributed Consensus** algorithm. In this scenario, multiple "Voter" agents independently decide on a value and send their decision to a "Leader" agent, who determines the final outcome based on a majority.

## Secure Private Inboxes

Communication in `HydraR` is handled through a `RestrictedState` wrapper, which provides each node with a private inbox. This ensures that nodes can only see messages explicitly sent to them, preventing state leakage and ensuring reliable multi-agent coordination.

## Loading the Workflow

We define the voters and the leader in a declarative YAML file. Each voter uses the same `voter_logic`, but operates as a unique node in the graph.

```{r setup_consensus}
library(HydraR)

# Load the consensus workflow from YAML
wf <- load_workflow("distributed_communication.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Architecture

We have three independent voters feeding their decisions into a single **Consensus Leader**.

```{r plot}
dag$plot()
```

## Running the Simulation

We execute the consensus run. Behind the scenes, the voters use `state$send_message()` to transmit their votes.

```{r execution}
# Execute the consensus run
final_run <- dag$run(wf$initial_state)

# View final result from the Leader
print(final_run$results$Leader$output$final_consensus)
print(final_run$results$Leader$output$vote_table)
```

## Verification and Auditing

`HydraR` supports an optional `MessageLog` that records all transmissions for later audit. This is critical for debugging complex distributed systems or verifying that agents are communicating as intended.

---
<!-- APAF Bioinformatics | distributed_communication.Rmd | Approved | 2026-04-03 -->
