---
title: "Gemini CLI Driver Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gemini CLI Driver Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(HydraR)
```

This vignette demonstrates how to use the `GeminiCLIDriver` to power an LLM node in `HydraR`. The CLI driver is ideal for local development and environments where direct API access might be restricted or where you want to leverage pre-configured CLI tools.

## Setup

Ensure you have the `gemini` CLI installed and authenticated on your system.

## Loading the Workflow

We define a simple "Poetic Bioinformatician" agent in a YAML file.

```{r load}
library(HydraR)

# Load the workflow from YAML
wf <- load_workflow("gemini_cli_demo.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Running with the CLI Driver

The `dag$run()` method will invoke the `GeminiCLIDriver` for the `writer` node.

```{r running, eval = FALSE}
# Execute with the initial topic: "DNA sequencing"
final <- dag$run(wf$initial_state)

# View the poetic output
cat(final$results$writer$output)
```

## Driver Specifics

The `GeminiCLIDriver` translates the agent's role and prompt into a standard CLI command (e.g., `gemini --role '...' --prompt '...'`). This allows for:
1.  **Observability**: You can see the raw CLI commands being executed in your terminal or logs.
2.  **Configuration**: CLI-specific options like `--model` or `--temperature` can be passed via the `cli_opts` parameter in the node definition.

---
<!-- APAF Bioinformatics | gemini_cli_demo.Rmd | Approved | 2026-04-03 -->
