---
title: "Hello World: Building an Agentic Loop"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello World: Building an Agentic Loop}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates how to construct a simple "Hello World" agentic loop using the `HydraR` framework. We will create a Directed Acyclic Graph (DAG) with two nodes: an **LLM Guesser** and a **Greeting Validator**.

## The Guesser-Validator Loop

1.  **LLM Guesser**: An agentically powered node that tries to guess a secret word ("hello").
2.  **Greeting Validator**: A deterministic R logic node that checks the guess. If the guess is incorrect, it triggers a loop back to the guesser.

## Declarative Workflow

We define the entire orchestration, including the looping logic and agent roles, in a YAML file.

```{r load}
library(HydraR)

# Load the agentic workflow from YAML
wf <- load_workflow("hello_world_agent.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Loop

The `plot()` method shows the iterative connection from the **Validator** back to the **Guesser**.

```{r plot}
dag$plot()
```

## Running the Agentic Loop

The system will continue to loop until the LLM produces the correct target greeting!

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY
final <- dag$run(wf$initial_state, max_steps = 10)

# View final results
cat("Final Guess:", final$state$get("Guesser"), "\n")
cat("Success Status:", final$results$Validator$output$valid, "\n")
```

## Summary

The "Hello World Agent" demonstrates the core power of `HydraR`: combining LLM reasoning with deterministic R code to create autonomous, self-correcting feedback loops. This pattern forms the foundation for more complex systems like software bug fixers and iterative researchers.

---
<!-- APAF Bioinformatics | hello_world_agent.Rmd | Approved | 2026-04-03 -->
