---
title: "Hello World: Your First DAG"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello World: Your First DAG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

This vignette provides the simplest "Hello World" entry point for `HydraR`, demonstrating the declarative **YAML + Mermaid** pattern for defining and running a Directed Acyclic Graph (DAG) of logic nodes.

## Setup

```{r setup}
library(HydraR)
```

## Loading the Workflow

Instead of writing R code to manually instantiate nodes and edges, we define the entire architecture in a separate YAML file. This separates the **orchestration** from the **execution**.

```{r load}
# Load the declarative workflow definition
wf <- load_workflow("hello_world.yml")

# Spawn the compiled DAG
dag <- spawn_dag(wf)
```

## Visualizing the Architecture

We can visualize the DAG directly from the compiled object.

```{r plot}
dag$plot()
```

## Running the DAG

We execute the DAG with the initial state provided in the YAML.

```{r run}
# Running without a checkpointer for this simple example
final <- dag$run(wf$initial_state)

# Inspect the final output from the 'process_data' node
print(final$results$process_data$output$processed_result)
```

## Summary

This pattern allows you to:
1.  **Visualize** your logic before running it.
2.  **Externalize** node logic into blocks that can be easily audited.
3.  **Scale** to more complex LLM-driven agents by simply changing the `type` in the Mermaid node tags.

---
<!-- APAF Bioinformatics | hello_world.Rmd | Approved | 2026-04-03 -->
