---
title: "Hierarchical Agents: Critique and Revise"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hierarchical Agents: Critique and Revise}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates the **Critique and Revise** pattern, a fundamental architectural building block for high-quality agentic systems. In this workflow, two LLM agents—a **Writer** and a **Reviewer**—collaborate to refine a creative story through an iterative feedback loop.

## The Feedback Mechanism

1.  **Story Author**: Drafts the initial story based on the prompt. If feedback is received, it revises the draft to address the editor's concerns.
2.  **Literary Editor**: Critically reviews the story using internal quality criteria. If the story is approved, the loop terminates. Otherwise, it provides specific, constructive critique back to the author.

## Declarative Loop Definition

The logic for both prompt generation and the looping condition is defined in a YAML file.

```{r load}
library(HydraR)

# Load the hierarchical workflow from YAML
wf <- load_workflow("hierarchical_agents.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Hierarchy

The `plot()` method reveals the recursive interaction between the two agents.

```{r plot}
dag$plot()
```

## Executing the Collaborative Run

The `run()` method will continue to iterate until the `Reviewer` node returns an "Approved" status or the `max_steps` limit is hit.

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY
final <- dag$run(wf$initial_state, max_steps = 10)

# Display the final story and editor's feedback
cat("Final Story Draft:\n", final$state$get("Writer"), "\n")
cat("Final Feedback:", final$state$get("Reviewer"), "\n")
```

## Resilience and Durability

For long creative sessions, you can attach a `Checkpointer` (e.g., `DuckDBSaver`). This allows you to pause the process and resume it later, ensuring that no creative progress is lost due to API timeouts or system interruptions.

---
<!-- APAF Bioinformatics | hierarchical_agents.Rmd | Approved | 2026-04-03 -->
