---
title: "LLM Fan-Out and Fan-In Synthesis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LLM Fan-Out and Fan-In Synthesis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates the **Fan-Out / Fan-In** pattern using `HydraR`. Instead of manual node configuration, we use a declarative YAML syntax to define complex multi-agent topologies.

## Collaborative Storytelling Pattern

We model a collaborative writing process where a **Director** sets the stage, multiple specialized **Writers** (Action, Mystery, Romance) independently generate plot lines in parallel, and a **Master Editor** synthesizes them into a cohesive story.

This pattern is ideal for:
1.  **Parallel Creativity**: Generating diverse perspectives on the same input.
2.  **Synthesis**: Combining isolated agent outputs into a final consensus.

## Declarative Orchestration

We define the architecture in `parallel_fan_out.yml`.

```{r setup}
library(HydraR)

# Load the workflow from YAML
wf <- load_workflow("parallel_fan_out.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Parallelism

The `plot()` method shows the one-to-many "Fan-Out" from the Director and the many-to-one "Fan-In" to the Editor.

```{r plot}
dag$plot()
```

## Running the Parallel Chain

When `dag$run()` is called, the Director node executes first. Then, the Three writer nodes are evaluated (potentially in parallel if `furrr` is configured). Finally, the Master Editor consolidates the results.

```{r running, eval = FALSE}
# Execute with the initial premise
final <- dag$run(wf$initial_state)

# View the final synthesized story
cat(final$state$get("editor"))
```

## Scaling with Worktrees

For heavy parallel processing (e.g., each writer needing its own git environment), you can set `use_worktrees = TRUE` in the `run()` method. This ensures that each "branch" of the Fan-Out has a completely isolated filesystem for execution, preventing race conditions.

---
<!-- APAF Bioinformatics | parallel_fan_out.Rmd | Approved | 2026-04-03 -->
