---
title: "Round-Trip Resiliency: Status-Colored Plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Round-Trip Resiliency: Status-Colored Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates how to use the `HydraR` status-colored plotting API. By defining a workflow in YAML and running it, you can generate a **status-colored** Mermaid visualization that highlights nodes based on their execution outcome.

## The Quality Control Loop

We simulate a "Quality Control" loop where initial findings from a search are checked. If the check fails, the graph routes execution back to a "Deep Search" node to gather more detail before re-summarizing.

```{r setup}
library(HydraR)

# Load the resilient loop from YAML
wf <- load_workflow("round_trip_resiliency.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Spec

We start with a clean graph.

```{r plot_spec}
dag$plot()
```

## Running the Scenario

In this example, the `Check Quality` node is hardcoded to fail on its first run and succeed on its second.

```{r run}
# Run with a max_steps limit to prevent infinite loops
final <- dag$run(wf$initial_state, max_steps = 10)

# Display the final state
print(final$results$Publish$output)
```

## Generating the Status Plot

After the execution, the `plot(status = TRUE)` method uses the internal `trace_log` to apply styles to the Mermaid graph:
-   **Green**: Successful execution.
-   **Red**: Failed execution.
-   **Yellow**: Paused or currently active.

```{r plot_status}
# Export status-colored Mermaid
dag$plot(status = TRUE)
```

## Summary

`HydraR` ensures that even complex, cyclic graphs can be audited and visualized effectively, making it easier to debug multi-agent coordination issues and trace the path taken through the consensus engine.

---
<!-- APAF Bioinformatics | round_trip_resiliency.Rmd | Approved | 2026-04-03 -->
