---
title: "Agentic Software Bug Assistant: Self-Healing Loops"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Software Bug Assistant: Self-Healing Loops}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(HydraR)
```

This vignette demonstrates the **Software Bug Assistant** pattern, a classic "Generate-and-Test" self-healing loop. In this scenario, two agent nodes work together to resolve a software bug by proposing patches and verifying them against a test suite.

## The Remediation Cycle

1.  **Debugger Agent**: Reviews the bug report and any previous test failures, then proposes a corrected R code snippet.
2.  **Test Suite**: A logic node that simulates running a test suite against the proposed patch. If the test fails, it provides the error trace back to the Debugger.

## Declarative Self-Healing

The entire loop, including the testing logic and repair prompts, is defined in a YAML file.

```{r load}
library(HydraR)

# Load the bug assistant workflow
wf <- load_workflow("software_bug_assistant.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Visualizing the Repair Loop

The graph shows the tight feedback loop between the **Debugger Agent** and the **Test Suite**.

```{r plot}
dag$plot()
```

## Running the Automatic Remediation

The DAG will continue to iterate, with the `Analyzer` refining its patch based on the `Tester`'s feedback, until the test suite passes (or the `max_steps` limit is reached).

```{r run, eval = FALSE}
# Note: Requires a valid GEMINI_API_KEY
final <- dag$run(wf$initial_state, max_steps = 10)

# Display the final patch and test outcome
cat("Final Patch Proposed:\n", final$state$get("Analyzer"), "\n")
cat("Test Results:", final$state$get("Tester")$test_feedback, "\n")
```

## Summary

The `Bug Assistant` pattern is a cornerstone of autonomous software engineering. By combining LLMs' reasoning with deterministic code execution (testing), `HydraR` enables the creation of robust, self-correcting agents that can maintain and repair codebases with minimal human intervention.

---
<!-- APAF Bioinformatics | software_bug_assistant.Rmd | Approved | 2026-04-03 -->
