---
title: "Isolated Execution with Git Worktrees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Isolated Execution with Git Worktrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(HydraR)
```

A major challenge in building autonomous agents is **isolation**. When an agent modifies a file (e.g., writing code or fixing a bug), it can easily cause conflicts with other parallel agents or your main working directory. `HydraR` solves this by natively integrating with **Git Worktrees**.

## Why Use Worktrees?

1.  **No Conflicts**: Parallel agents can modify the same files (like `DESCRIPTION`) in separate worktrees without corrupted git states.
2.  **Explicit Merging**: Changes only appear in your main directory when a **Merge Harmonizer** node explicitly merges them back to the primary branch.
3.  **Hardened Context**: If an agent runs a destructive command, it only affects its temporary worktree, leaving your primary codebase safe.

## Setting Up a Test Repository

To avoid modifying your actual project, let's create a temporary Git repository.

```{r setup, eval = FALSE}
library(HydraR)

# Create a temporary folder
repo_root <- file.path(tempdir(), "hydra-toy-repo")
dir.create(repo_root)

# Initialize Git
withr::with_dir(repo_root, {
  system("git init -b main")
  system("git config user.email 'apaf@example.com'")
  system("git config user.name 'APAF Agent'")
  writeLines("# Toy Project", "README.md")
  system("git add README.md")
  system("git commit -m 'Initial commit'")
})
```

## Declarative Isolation

We define the workflow with a **Code Creator** and a **Merge Harmonizer**.

```{r load}
# Load the isolated worktree workflow
wf <- load_workflow("worktree_isolation.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Running with Worktree Isolation

By setting `use_worktrees = TRUE`, `HydraR` will:
1.  **Spawn a Worktree**: Create a new git branch and directory for each node.
2.  **Execute Isolated**: The LLM agent operates in the worktree directory.
3.  **Harmonize**: The `Merge Harmonizer` node consolidates the results.

```{r run, eval = FALSE}
# Run with worktree isolation enabled
final <- dag$run(
  initial_state = wf$initial_state,
  use_worktrees = TRUE,
  repo_root = repo_root,
  fail_if_dirty = FALSE
)

# Verify the file was merged into the main repo branch
list.files(repo_root)
# "README.md" "math_fun.R"
```

## Summary

Git Worktrees provide a robust, low-overhead mechanism for isolating agent activities. `HydraR` abstracts this complexity, allowing you to build parallel, safe multi-agent networks that respect your primary codebase's integrity.

---
<!-- APAF Bioinformatics | worktree_isolation.Rmd | Approved | 2026-04-03 -->
