---
title: "LLM Fan-Out and Fan-In Synthesis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LLM Fan-Out and Fan-In Synthesis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette demonstrates how to implement a **Fan-Out / Fan-In** pattern using the `GeminiCLIDriver` to power agentic nodes in `HydraR`.

Instead of writing R code to manually configure each node, we will use `HydraR`'s declarative YAML syntax. This makes defining complex multi-agent network topologies clean and portable.

## The Scenario

Imagine we are writing a collaborative story. The process requires a **Fan-Out** to multiple specialized writers (e.g., Action, Mystery, Romance) who independently generate distinct plot lines based on a central premise. Afterward, a **Fan-In** occurs where a Master Editor synthesizes these independent plots into a single, cohesive narrative.

This pattern is highly useful when:
1.  You need parallel brainstorming or multi-perspective analysis on the same input data.
2.  You want to synthesize diverse, isolated outputs into a final consensus or summary.

## Setup

```{r setup}
library(HydraR)
```

## The Declarative Definition (`fan_out_fan_in.yml`)

We define the entire orchestration architecture in a separate `fan_out_fan_in.yml` file. This file contains:
1.  **Graph**: The Mermaid string dictating the connections and node parameters (`type=llm`).
2.  **Roles**: The system prompts for the LLM models.
3.  **Logic**: Inline R snippets for building prompts dynamically from state.
4.  **Initial State**: The starting variables for our workflow.

Here is the content of `fan_out_fan_in.yml`:

```yaml
graph: |
  graph TD
    director["Director | type=llm | role_id=director | driver=gemini | prompt_id=prompt_director"]
    writer_action["Action Writer | type=llm | role_id=writer_action | driver=gemini | prompt_id=prompt_action"]
    writer_mystery["Mystery Writer | type=llm | role_id=writer_mystery | driver=gemini | prompt_id=prompt_mystery"]
    writer_romance["Romance Writer | type=llm | role_id=writer_romance | driver=gemini | prompt_id=prompt_romance"]
    editor["Master Editor | type=llm | role_id=editor | driver=gemini | prompt_id=prompt_editor"]

    director --> writer_action
    director --> writer_mystery
    director --> writer_romance

    writer_action --> editor
    writer_mystery --> editor
    writer_romance --> editor

roles:
  director: "You are the creative director setting the stage."
  writer_action: "You are an action-thriller writer. Write a short, fast-paced scene."
  writer_mystery: "You are a mystery writer. Write a short, suspenseful scene."
  writer_romance: "You are a romance writer. Write a short, emotional scene."
  editor: "You are the master editor. Combine the scenes into a cohesive short story."

logic:
  prompt_director: >
    sprintf("Expand slightly on this premise: %s", state$get("premise"))
  prompt_action: >
    sprintf("Based on the director's vision: %s\nWrite an action scene.", state$get("director"))
  prompt_mystery: >
    sprintf("Based on the director's vision: %s\nWrite a mystery scene.", state$get("director"))
  prompt_romance: >
    sprintf("Based on the director's vision: %s\nWrite a romantic or emotional scene.", state$get("director"))
  prompt_editor: >
    sprintf(
      "Synthesize these three scenes into one short story:\n\nAction: %s\n\nMystery: %s\n\nRomance: %s",
      state$get("writer_action"),
      state$get("writer_mystery"),
      state$get("writer_romance")
    )

start_node: director

initial_state:
  premise: "A cybernetic cat discovers a hidden underground city."
```

## Loading and Compiling the DAG

We load the YAML file and spawn the DAG using the automatic node factory, which automatically resolves `type=llm` annotations and retrieves logic/roles from the registries.

```{r load_wf}
# 1. Load the workflow definitions into the registries
wf <- load_workflow("fan_out_fan_in.yml")

# 2. Compile the Directed Acyclic Graph
dag <- spawn_dag(wf, auto_node_factory())
```

## Executing the DAG

Running the DAG will automatically follow the Mermaid structure. The `director` node will execute first. Then, depending on execution context, the three writer nodes will execute. Finally, the `editor` node will run, using the output accumulated in the state from the writers.

```{r running}
# Execute with the initial premise extracted from the YAML file
res <- dag$run(initial_state = wf$initial_state)

# View the final synthesized story
cat(res$results$editor$output)
```

## Visualizing the Flow

We can inspect the execution path and configuration right from R.

```{r, results='asis'}
cat("```mermaid\n")
cat(dag$plot(type = "mermaid", details = TRUE))
cat("\n```\n")
```

## Conclusion

The Fan-Out / Fan-In pattern is a robust way to generate diverse perspectives or sub-components of a task in parallel before synthesizing them. Using `HydraR`'s declarative YAML framework, complex multi-agent topologies become easily portable and readable configuration files instead of lengthy boilerplate code.

<!-- APAF Bioinformatics | fan_out_fan_in.Rmd | Approved | 2026-03-31 -->
