---
title: "DuckDB Restart Reprex: Automated Workflow Persistence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DuckDB Restart Reprex: Automated Workflow Persistence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates a **Minimal Reproducible Example (Reprex)** for the `HydraR` core feature: **Automated State Persistence & Restart**. 

In high-complexity multi-agent workflows, code can fail due to environmental factors, API timeouts, or logic errors. `HydraR` allows you to:
1.  **Checkpoint** every step of a DAG to a `DuckDB` database.
2.  **Pause** execution when a step encounters a problem.
3.  **Resume** exactly from the paused node after the state has been "fixed."

## Setup

Load the library and define a temporary DuckDB path.

```{r lib_load}
library(HydraR)
db_path <- tempfile(fileext = ".duckdb")
```

## 1. Defining the Logic

We register three logic functions globally. Note how `check_fixed` inspects the state for a `fixed` flag — if it's not `TRUE`, the node **pauses** execution instead of crashing.

```{r register_logic}
register_logic("init_proc", function(state) {
  list(status = "success", output = "Initialization complete.")
})

register_logic("check_fixed", function(state) {
  if (!isTRUE(state$get("fixed"))) {
    # Return "pause" status — HydraR saves state and stops cleanly
    return(list(status = "pause", output = "Waiting for manual fix."))
  }
  list(status = "success", output = "System recovered!")
})

register_logic("finalize_proc", function(state) {
  list(status = "success", output = "All steps finished.")
})
```

## 2. Building the DAG Inline

We define a simple 3-step pipeline using Mermaid syntax.

```{r build_dag}
mermaid_src <- '
graph TD
  Step1["Initialization | type=logic | logic_id=init_proc"]
  Step2["Risky Logic | type=logic | logic_id=check_fixed"]
  Step3["Conclusion | type=logic | logic_id=finalize_proc"]
  Step1 --> Step2
  Step2 --> Step3
'

dag <- mermaid_to_dag(mermaid_src)
dag$set_start_node("Step1")
```

## 3. Initial Run (The Pause)

We run the DAG with `fixed = FALSE`. Step2 will **pause** the pipeline and the checkpointer saves the state to DuckDB.

```{r run_1}
# Configure DuckDB Persistence
saver <- DuckDBSaver$new(db_path = db_path)
tid <- "reprex-session-001"

# Run 1: Expected to pause at Step2
res1 <- dag$run(
  thread_id = tid,
  checkpointer = saver,
  initial_state = list(fixed = FALSE, counter = 0)
)

cat("Execution Status:", res1$status, "\n")
cat("Paused at:", res1$paused_at, "\n")
```

## 4. Inspecting the Checkpoint

Even though the execution paused, the `thread_id` and the state (including Step1's result) are safely stored in DuckDB.

```{r inspect}
con <- DBI::dbConnect(duckdb::duckdb(), db_path)
DBI::dbGetQuery(con, "SELECT thread_id, updated_at FROM agent_checkpoints")
DBI::dbDisconnect(con, shutdown = TRUE)
```

## 5. The Restart (The Success)

Now we "fix" the state and restart from `Step2`. `HydraR` detects the existing checkpoint, merges our fix (`fixed = TRUE`) into the restored state, and re-executes from `Step2` onwards — **skipping** `Step1` entirely.

```{r run_2}
# Fixing the state: Set 'fixed' to TRUE
new_state <- list(fixed = TRUE)

# Run 2: Resume from Step2 using the SAME thread_id
final_results <- dag$run(
  thread_id = tid,
  checkpointer = saver,
  initial_state = new_state,
  resume_from = "Step2"
)

# Verify results
cat("Final Status:", final_results$status, "\n")
cat("Step2 Output:", final_results$results$Step2$output, "\n")
cat("Step3 Output:", final_results$results$Step3$output, "\n")
```

## Summary

The `DuckDB` checkpointer ensures that no computational progress is lost. This is critical for agentic workflows involving expensive LLM calls or long-running bioinformatics pipelines.

---
<!-- APAF Bioinformatics | reprex_duckdb.Rmd | Approved | 2026-04-01 -->
