---
title: "Agentic Academic Researcher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Academic Researcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Academic Research** pattern using `HydraR`. 
We define a linear, multi-stage agentic workflow designed to synthesize academic literature into a final report using the **Gemini CLI**.

The pipeline comprises three agents:
1. **Searcher Node**: Given a research topic, finds relevant academic papers.
2. **Summarizer Node**: Ingests the raw papers and extracts key findings.
3. **Compiler Node**: Takes the extracted findings and formats them into a comprehensive literature review.

## Setup

```{r}
library(HydraR)

# Initialize the Gemini CLI driver
# Note: This assumes the 'gemini' CLI is installed and configured on your system.
driver <- GeminiCLIDriver$new()
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Searcher Node

Utilizing an LLM to identify key papers for a topic.

```{r}
searcher_node <- AgentLLMNode$new(
  id = "Searcher",
  label = "Literature Searcher",
  role = "You are an academic research assistant. Identify 2-3 key hypothetical papers for the given topic. Output the result as a simple list.",
  driver = driver,
  prompt_builder = function(state) {
    sprintf("Research Topic: %s\nProvide a list of 2-3 paper titles and brief descriptions.", state$get("research_topic"))
  }
)

dag$add_node(searcher_node)
```

### 2. The Summarizer Node

Extracts value from the search results.

```{r}
summarizer_node <- AgentLLMNode$new(
  id = "Summarizer",
  label = "Content Summarizer",
  role = "You are a scientific editor. Summarize the following paper list into key highlights.",
  driver = driver,
  prompt_builder = function(state) {
    sprintf("Paper List: %s", state$get("Searcher"))
  }
)

dag$add_node(summarizer_node)
```

### 3. The Compiler Node

Drafts the final markdown report.

```{r}
compiler_node <- AgentLLMNode$new(
  id = "Compiler",
  label = "Report Compiler",
  role = "You are a technical writer. Format the following summaries into a professional markdown report with headers.",
  driver = driver,
  prompt_builder = function(state) {
    sprintf("Topic: %s\nSummaries: %s", state$get("research_topic"), state$get("Summarizer"))
  }
)

dag$add_node(compiler_node)
```

## Defining Transitions

Configure a straight-through pipeline: `Searcher -> Summarizer -> Compiler`

```{r}
dag$set_start_node("Searcher")

dag$add_edge("Searcher", "Summarizer")
dag$add_edge("Summarizer", "Compiler")

compiled_dag <- dag$compile()
```

## Visualizing the Workflow

```{r, results = 'asis'}
cat("```mermaid\n")
cat(compiled_dag$plot(type = "mermaid"))
cat("\n```\n")
```

## Running the Scenario

Provide the topic and run the pipeline.

```{r, eval = FALSE}
initial_state <- list(
  research_topic = "CRISPR Gene Editing"
)

cat("Starting Literature Pipeline...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 5)

cat("\n--- PIPELINE EXECUTION COMPLETE ---\n")
cat(result$state$get("Compiler"), "\n")
```

The system successfully routed the data through all three agents to build a consolidated report!

<!-- APAF Bioinformatics | academic_research.Rmd | Approved | 2026-03-29 -->
