---
title: "Agentic Academic Researcher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Academic Researcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(HydraR)
```

## Introduction

This vignette demonstrates the **Academic Research** pattern using `HydraR`. 
We define a linear, multi-stage agentic workflow designed to synthesize academic literature into a final report using the **Gemini CLI**.

The pipeline comprises three agents:
1. **Searcher Node**: Given a research topic, finds relevant academic papers.
2. **Summarizer Node**: Ingests the raw papers and extracts key findings.
3. **Compiler Node**: Takes the extracted findings and formats them into a comprehensive literature review.

## Setup

```{r library_load}
library(HydraR)
```

## Checkpointer Configuration

HydraR supports optional state checkpointing after each node execution.
Choose from four modes:

- **`"none"`** -- Ephemeral run, no persistence (default).
- **`"memory"`** -- In-memory via `MemorySaver` (lost on session exit).
- **`"rds"`** -- File-based via `RDSSaver` (lightweight, base R only).
- **`"duckdb"`** -- Database-backed via `DuckDBSaver` (requires `duckdb` package).

```{r}
# ── USER CONFIGURATION ──────────────────────────────────────────
CHECKPOINTER_MODE <- "none" # "none" | "memory" | "rds" | "duckdb"
RDS_DIR <- "checkpoints"
DUCKDB_PATH <- "checkpoints/academic_research.duckdb"
THREAD_ID <- "academic_research_run_1"
# ────────────────────────────────────────────────────────────────

checkpointer <- switch(CHECKPOINTER_MODE,
  "memory" = MemorySaver$new(),
  "rds" = RDSSaver$new(dir = RDS_DIR),
  "duckdb" = {
    if (requireNamespace("duckdb", quietly = TRUE)) {
      if (!dir.exists(dirname(DUCKDB_PATH))) dir.create(dirname(DUCKDB_PATH), recursive = TRUE)
      DuckDBSaver$new(db_path = DUCKDB_PATH)
    } else {
      warning("duckdb package is required but not installed. Falling back to none.")
      NULL
    }
  },
  "none" = NULL,
  stop(sprintf("Invalid CHECKPOINTER_MODE: '%s'", CHECKPOINTER_MODE))
)

thread_id <- if (!is.null(checkpointer)) THREAD_ID else NULL
```

## Defining the Workflow Components

To keep our architecture clean, we store all workflow components—initial configuration, LLM prompts, and agent roles—in a central registry.

```{r logic_registry}
research_logic_registry <- list(
  # 0. Initial Research Topic
  initial_state = list(
    research_topic = "CRISPR Gene Editing"
  ),

  # 1. Agent Roles
  roles = list(
    Searcher = "You are an academic research assistant. Identify 2-3 key hypothetical papers for the given topic. Output the result as a simple list.",
    Summarizer = "You are a scientific editor. Summarize the following paper list into key highlights.",
    Compiler = "You are a technical writer. Format the following summaries into a professional markdown report with headers."
  ),

  # 2. Prompt Builders
  prompts = list(
    Searcher = function(state) {
      sprintf("Research Topic: %s\nProvide a list of 2-3 paper titles and brief descriptions.", state$get("research_topic"))
    },
    Summarizer = function(state) {
      sprintf("Paper List: %s", state$get("Searcher"))
    },
    Compiler = function(state) {
      sprintf("Topic: %s\nSummaries: %s", state$get("research_topic"), state$get("Summarizer"))
    }
  )
)
```

## The Node Factory

We use a factory function to dynamically create nodes based on parameters defined in the Mermaid graph.

```{r factory}
research_node_factory <- function(id, label, params) {
  # Driver resolution from Mermaid params
  driver_obj <- if (params$driver == "gemini") GeminiCLIDriver$new() else NULL

  AgentLLMNode$new(
    id = id,
    label = label,
    role = research_logic_registry$roles[[id]],
    driver = driver_obj,
    prompt_builder = research_logic_registry$prompts[[id]]
  )
}
```

## Building the DAG via Mermaid

We define the entire workflow architecture as a Mermaid string. This string serves as the single source of truth for both structure and node metadata.

```{r mermaid_source}
mermaid_graph <- "
graph TD
  Searcher[Literature Searcher | driver=gemini] --> Summarizer
  Summarizer[Content Summarizer | driver=gemini] --> Compiler
  Compiler[Report Compiler | driver=gemini]
"

# Instantiate the DAG
dag <- AgentDAG$from_mermaid(mermaid_graph, node_factory = research_node_factory)
compiled_dag <- dag$compile()
```

## Visualizing the Workflow

```{r, results = 'asis'}
cat("```mermaid\n")
cat(compiled_dag$plot(type = "mermaid"))
cat("\n```\n")
```

## Running the Scenario

Provide the topic and run the pipeline.

```{r, eval = FALSE}
cat(sprintf("Starting Literature Pipeline (checkpointer: %s)...\n", CHECKPOINTER_MODE))

result <- compiled_dag$run(
  initial_state = research_logic_registry$initial_state,
  max_steps = 5,
  checkpointer = checkpointer,
  thread_id = thread_id
)

cat("\n--- PIPELINE EXECUTION COMPLETE ---\n")
cat(result$state$get("Compiler"), "\n")
```

The system successfully routed the data through all three agents to build a consolidated report!

<!-- APAF Bioinformatics | academic_research.Rmd | Approved | 2026-03-29 -->
