---
title: "Agentic Academic Researcher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Academic Researcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Academic Research** pattern using `HydraR`. 
We define a linear, multi-stage agentic workflow designed to synthesize academic literature into a final report.

The pipeline comprises three agents:
1. **Searcher Node**: Given a research topic, finds relevant academic papers.
2. **Summarizer Node**: Ingests the raw papers and extracts key findings.
3. **Compiler Node**: Takes the extracted findings and formats them into a comprehensive literature review.

This pattern is a perfect example of a **sequential processing pipeline** without cyclic loops, where each node enriches the state.

## Setup

```{r}
library(HydraR)
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Searcher Node

Mocking an LLM utilizing a tool to search PubMed or ArXiv.

```{r}
searcher_node <- AgentLogicNode$new(id = "Searcher", logic_fn = function(state, memory = NULL) {
  topic <- state$get("research_topic")

  # Mock search results
  papers <- list(
    list(title = paste(topic, "Basics"), content = "Introduction to the topic."),
    list(title = paste("Advanced", topic), content = "Complex analysis of the topic.")
  )

  list(
    status = "SUCCESS",
    output = list(
      raw_papers = papers,
      search_status = "Found 2 papers."
    )
  )
})

dag$add_node(searcher_node)
```

### 2. The Summarizer Node

Extracts value from the search results.

```{r}
summarizer_node <- AgentLogicNode$new(id = "Summarizer", logic_fn = function(state, memory = NULL) {
  papers <- state$get("raw_papers")

  # Mocking summarization
  summaries <- lapply(papers, function(p) {
    sprintf("- **%s**: Highlights - %s", p$title, p$content)
  })

  list(
    status = "SUCCESS",
    output = list(
      paper_summaries = unlist(summaries)
    )
  )
})

dag$add_node(summarizer_node)
```

### 3. The Compiler Node

Drafts the final markdown report.

```{r}
compiler_node <- AgentLogicNode$new(id = "Compiler", logic_fn = function(state, memory = NULL) {
  topic <- state$get("research_topic")
  summaries <- state$get("paper_summaries")

  report <- paste0(
    "# Literature Review: ", topic, "\n\n",
    "## Key Findings\n",
    paste(summaries, collapse = "\n")
  )

  list(
    status = "SUCCESS",
    output = list(
      final_report = report
    )
  )
})

dag$add_node(compiler_node)
```

## Defining Transitions

Configure a straight-through pipeline: `Searcher -> Summarizer -> Compiler`

```{r}
dag$set_start_node("Searcher")

dag$add_edge("Searcher", "Summarizer")
dag$add_edge("Summarizer", "Compiler")

compiled_dag <- dag$compile()
```

## Running the Scenario

Provide the topic and run the pipeline.

```{r}
initial_state <- list(
  research_topic = "CRISPR Gene Editing"
)

cat("Starting Literature Pipeline...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 5)

cat("\n--- PIPELINE EXECUTION COMPLETE ---\n")
cat("Search Status:", result$state$get("search_status"), "\n\n")
cat(result$state$get("final_report"), "\n")
```

The system successfully routed the data through all three agents to build a consolidated report!

<!-- APAF Bioinformatics | academic_research.Rmd | Approved | 2026-03-29 -->
