---
title: "Agentic Blog Writer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Blog Writer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Blog Writer** pattern using `HydraR`. 

Content creation often involves a multi-stage pipeline consisting of brainstorming, drafting, and reviewing. We model this as a DAG with a mix of linear execution and cyclic feedback.

1. **Outliner Node**: Generates a structured outline based on a topic.
2. **Drafter Node**: Takes the outline and writes the draft. It also reacts to feedback from the Editor.
3. **Editor Node**: Reviews the draft for SEO and readability. If it fails the checks, it loops back to the Drafter.

## Setup

```{r}
library(HydraR)
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Outliner Node

This node only runs once at the beginning of the pipeline.

```{r}
outliner_node <- AgentLogicNode$new(id = "Outliner", logic_fn = function(state, memory = NULL) {
  topic <- state$get("blog_topic")

  list(
    status = "SUCCESS",
    output = list(
      outline = sprintf("1. Introduction matching '%s'\n2. Detailed Body\n3. Conclusion", topic)
    )
  )
})

dag$add_node(outliner_node)
```

### 2. The Drafter Node

This node merges the outline with any editor feedback.

```{r}
drafter_node <- AgentLogicNode$new(id = "Drafter", logic_fn = function(state, memory = NULL) {
  outline <- state$get("outline")
  feedback <- state$get("editor_feedback")

  attempts <- state$get("draft_attempts")
  if (is.null(attempts)) attempts <- 0
  attempts <- attempts + 1

  # Mocking drafting: on the second attempt, we "fix" the SEO issue
  if (attempts == 1) {
    draft <- "Here is a blog post based on the outline. It is very generic."
  } else {
    draft <- "Here is an SEO-optimized blog post with correct keywords and structure."
  }

  list(
    status = "SUCCESS",
    output = list(
      blog_draft = draft,
      draft_attempts = attempts
    )
  )
})

dag$add_node(drafter_node)
```

### 3. The Editor Node

Checks the draft for specific keywords.

```{r}
editor_node <- AgentLogicNode$new(id = "Editor", logic_fn = function(state, memory = NULL) {
  draft <- state$get("blog_draft")

  if (grepl("SEO-optimized", draft)) {
    list(
      status = "SUCCESS",
      output = list(
        is_published = TRUE,
        editor_feedback = "Approved for publishing."
      )
    )
  } else {
    list(
      status = "SUCCESS",
      output = list(
        is_published = FALSE,
        editor_feedback = "The draft lacks SEO optimization. Please revise."
      )
    )
  }
})

dag$add_node(editor_node)
```

## Defining Transitions

Combine linear and cyclic edges.

```{r}
dag$set_start_node("Outliner")

dag$add_edge("Outliner", "Drafter")
dag$add_edge("Drafter", "Editor")

dag$add_conditional_edge(
  from = "Editor",
  test = function(out) {
    isTRUE(out$is_published)
  },
  if_true = NULL, # Stop!
  if_false = "Drafter" # Back to drafter
)

compiled_dag <- dag$compile()
```

## Running the Scenario

```{r}
initial_state <- list(
  blog_topic = "Agentic Workflows in R"
)

cat("Starting Blog Creation Engine...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 10)

cat("\n--- BLOG PUBLICATION STATUS ---\n")
cat("Outline Created:\n", result$state$get("outline"), "\n")
cat("Total Drafts:", result$state$get("draft_attempts"), "\n")
cat("Final Output:\n", result$state$get("blog_draft"), "\n")
cat("Editor Decision:", result$state$get("editor_feedback"), "\n")
```

The DAG easily handles complex flows involving both straight-through processing and localized feedback loops.

<!-- APAF Bioinformatics | blog_writer.Rmd | Approved | 2026-03-29 -->
