---
title: "Agentic Data Science Assistant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Data Science Assistant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Data Science Assistant** pattern using `HydraR`. 

In machine learning workflows, hyperparameter tuning is an inherently iterative process. We can map this paradigm to an `AgentDAG`. 
The workflow contains an initial data setup phase followed by a cyclic model optimization phase:
1. **Data Cleaner Node**: Ingests and prepares the raw data.
2. **Model Trainer Node**: Mocks training a model using a set of hyperparameters.
3. **Evaluator Node**: Assesses the model's accuracy. If it falls below the target threshold, it routes back to the Trainer to adjust parameters.

## Setup

```{r}
library(HydraR)
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Data Cleaner Node

Prepares the data exactly once.

```{r}
cleaner_node <- AgentLogicNode$new(id = "DataCleaner", logic_fn = function(state, memory = NULL) {
    dataset <- state$get("raw_data")
    
    # Mock cleaning
    clean_data <- paste("Cleaned", dataset)
    
    list(
        status = "SUCCESS",
        output = list(
            clean_data = clean_data
        )
    )
})

dag$add_node(cleaner_node)
```

### 2. The Model Trainer Node

Trains the model iteratively. It increases the "epoch" or "complexity" every time it gets called by the Evaluator.

```{r}
trainer_node <- AgentLogicNode$new(id = "ModelTrainer", logic_fn = function(state, memory = NULL) {
    clean_data <- state$get("clean_data")
    
    complexity <- state$get("model_complexity")
    if (is.null(complexity)) complexity <- 1
    
    # Increase complexity on each training loop
    new_complexity <- complexity + 1
    
    # Mocking training: accuracy improves with complexity, capping at 0.95
    model_accuracy <- min(0.60 + (new_complexity * 0.10), 0.95)
    
    list(
        status = "SUCCESS",
        output = list(
            model_complexity = new_complexity,
            model_accuracy = model_accuracy,
            model_artifact = sprintf("Model_v%d", new_complexity)
        )
    )
})

dag$add_node(trainer_node)
```

### 3. The Evaluator Node

Compares the accuracy to the user's requirement.

```{r}
evaluator_node <- AgentLogicNode$new(id = "Evaluator", logic_fn = function(state, memory = NULL) {
    accuracy <- state$get("model_accuracy")
    target <- state$get("target_accuracy")
    
    if (accuracy >= target) {
        list(
            status = "SUCCESS",
            output = list(
                optimization_complete = TRUE,
                eval_message = sprintf("Target reached: %.2f >= %.2f", accuracy, target)
            )
        )
    } else {
        list(
            status = "SUCCESS",
            output = list(
                optimization_complete = FALSE,
                eval_message = sprintf("Accuracy %.2f is below %.2f. Retraining.", accuracy, target)
            )
        )
    }
})

dag$add_node(evaluator_node)
```

## Defining Transitions

Configure the linear start and the cyclic optimization loop.

```{r}
dag$set_start_node("DataCleaner")

dag$add_edge("DataCleaner", "ModelTrainer")
dag$add_edge("ModelTrainer", "Evaluator")

dag$add_conditional_edge(
    from = "Evaluator",
    test = function(out) {
        isTRUE(out$optimization_complete)
    },
    if_true = NULL,            # Done!
    if_false = "ModelTrainer"  # Needs more complexity
)

compiled_dag <- dag$compile()
```

## Running the Scenario

Let's test it with a target accuracy constraint.

```{r}
initial_state <- list(
    raw_data = "Titanic_Dataset.csv",
    target_accuracy = 0.85
)

cat("Starting AutoML Pipeline...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 10)

cat("\n--- TRAINING PIPELINE COMPLETE ---\n")
cat("Clean Data Used:", result$state$get("clean_data"), "\n")
cat("Final Model Artifact:", result$state$get("model_artifact"), "\n")
cat("Final Accuracy:", result$state$get("model_accuracy"), "\n")
cat("Evaluation Message:", result$state$get("eval_message"), "\n")
```

The DAG seamlessly executed the data prep before dropping into an iterative optimization loop until the target metric was satisfied!

<!-- APAF Bioinformatics | data_science.Rmd | Approved | 2026-03-29 -->
