---
title: "Agentic Checkpointing and Restarts with DuckDB"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Checkpointing and Restarts with DuckDB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(HydraR)
```

## Introduction

In the **Journal of Open Source Software (JOSS)**, computational reproducibility is paramount. `HydraR` enables robust scientific pipelines that can pause, halt on errors, and be cleanly restarted without losing context.

This vignette provides a **minimal reproducible example (reprex)** of how to use `DuckDBSaver` to persist agent state, simulate a failure, fix the underlying logic, and restart the execution seamlessly from the point of failure.

## Setup the Checkpointer

We initialize a `DuckDBSaver` to persist state across execution boundaries.

```{r checkpointer_setup, eval=FALSE}
library(HydraR)
library(duckdb)

db_path <- tempfile(fileext = ".duckdb")
checkpointer <- DuckDBSaver$new(db_path = db_path)
thread_id <- "reprex_run_001"
```

## Define the Workflow

We define a simple two-node workflow:
1. **Node A (Data Prep)**: Simulates an expensive operation.
2. **Node B (Analysis)**: Has an intentional bug that we will "fix" mid-flight.

```{r logic_nodes, eval=FALSE}
# Node A: A successful, expensive initial step
node_a <- AgentLogicNode$new(
  id = "Data_Prep",
  logic_fn = function(state) {
    message("Running expensive data preparation...")
    # Simulate work
    Sys.sleep(1)
    list(status = "SUCCESS", output = list(prep_done = TRUE, data = c(1, 2, 3)))
  }
)

# Node B: Initially buggy analysis step
node_b_buggy <- AgentLogicNode$new(
  id = "Analysis",
  logic_fn = function(state) {
    data <- state$get("Data_Prep")$data
    message("Running analysis...")
    # INTENTIONAL BUG: trying to sum a NULL value if data wasn't handled correctly
    stop("Simulated crash: unexpected data format in Analysis!")
  }
)

dag <- AgentDAG$new()
dag$add_node(node_a)
dag$add_node(node_b_buggy)
dag$add_edge("Data_Prep", "Analysis")
dag$compile()
```

## First Run: Catching the Failure

We run the DAG. It will persist the state after `Data_Prep` succeeds, but crash at `Analysis`.

```{r first_run, eval=FALSE, error=TRUE}
tryCatch({
  dag$run(
    initial_state = list(input = "Start"),
    checkpointer = checkpointer,
    thread_id = thread_id
  )
}, error = function(e) {
  message("DAG execution failed as expected: ", e$message)
})
```

At this point, the state after `Data_Prep` is safely stored in DuckDB.
We can retrieve it to inspect the context at the time of failure.

```{r inspect_state, eval=FALSE}
saved_state <- checkpointer$get(thread_id)
print(saved_state$get("Data_Prep"))
# $prep_done
# [1] TRUE
# $data
# [1] 1 2 3
```

## The Fix & Restart

Now, we "fix" the bug in Node B and create a new DAG. Instead of running the expensive `Data_Prep` again, we reload the saved state and instruct the DAG to resume from the `Analysis` node.

```{r restart, eval=FALSE}
# Fix the logic
node_b_fixed <- AgentLogicNode$new(
  id = "Analysis",
  logic_fn = function(state) {
    data <- state$get("Data_Prep")$data
    message("Running FIXED analysis...")
    list(status = "SUCCESS", output = list(result = sum(data)))
  }
)

# Rebuild DAG with fixed node
fixed_dag <- AgentDAG$new()
fixed_dag$add_node(node_a)
fixed_dag$add_node(node_b_fixed)
fixed_dag$add_edge("Data_Prep", "Analysis")
fixed_dag$compile()

# Restart from DuckDB checkpoint
message("Restarting from checkpoint...")
restarted_result <- fixed_dag$run(
  initial_state = checkpointer$get(thread_id),
  resume_from = "Analysis",
  checkpointer = checkpointer,
  thread_id = thread_id
)

message("Final Result: ", restarted_result$state$get("Analysis")$result)
# Output should be: Final Result: 6
```

## Conclusion

With a few lines of code, `HydraR` combined with `DuckDB` guarantees that long-running scientific and LLM workflows can survive crashes and be interactively debugged, saving compute resources and frustration.
