---
title: "LLM Fan-Out and Fan-In Synthesis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LLM Fan-Out and Fan-In Synthesis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette demonstrates how to implement a **Fan-Out / Fan-In** pattern using the `GeminiCLIDriver` to power agentic nodes in `HydraR`.

## The Scenario

Imagine we are writing a collaborative story. The process requires a **Fan-Out** to multiple specialized writers (e.g., Action, Mystery, Romance) who independently generate distinct plot lines based on a central premise. Afterward, a **Fan-In** occurs where a Master Editor synthesizes these independent plots into a single, cohesive narrative.

This pattern is highly useful when:
1.  You need parallel brainstorming or multi-perspective analysis on the same input data.
2.  You want to synthesize diverse, isolated outputs into a final consensus or summary.

## Setup

```{r setup}
library(HydraR)
```

## Defining the Workflow Components

To keep our architecture clean, we store all workflow componentsā€”initial configuration, LLM prompts, and agent rolesā€”in a central registry.

```{r logic_registry}
story_logic_registry <- list(
  # 0. Initial Configuration
  initial_state = list(
    premise = "A cybernetic cat discovers a hidden underground city."
  ),

  # 1. Agent Roles
  roles = list(
    director = "You are the creative director setting the stage.",
    writer_action = "You are an action-thriller writer. Write a short, fast-paced scene.",
    writer_mystery = "You are a mystery writer. Write a short, suspenseful scene.",
    writer_romance = "You are a romance writer. Write a short, emotional scene.",
    editor = "You are the master editor. Combine the scenes into a cohesive short story."
  ),

  # 2. Prompt Builders
  prompts = list(
    director = function(state) {
      sprintf("Expand slightly on this premise: %s", state$get("premise"))
    },
    writer_action = function(state) {
      sprintf("Based on the director's vision: %s\nWrite an action scene.", state$get("director"))
    },
    writer_mystery = function(state) {
      sprintf("Based on the director's vision: %s\nWrite a mystery scene.", state$get("director"))
    },
    writer_romance = function(state) {
      sprintf("Based on the director's vision: %s\nWrite a romantic or emotional scene.", state$get("director"))
    },
    editor = function(state) {
      # The editor accesses the outputs of all three writers
      sprintf(
        "Synthesize these three scenes into one short story:\n\nAction: %s\n\nMystery: %s\n\nRomance: %s",
        state$get("writer_action"),
        state$get("writer_mystery"),
        state$get("writer_romance")
      )
    }
  )
)
```

## The Node Factory

We use a factory function to dynamically resolve nodes and their drivers based on parameters defined in the Mermaid graph.

```{r factory}
story_node_factory <- function(id, label, params) {
  # Driver resolution from Mermaid params
  driver_obj <- if (!is.null(params$driver) && params$driver == "gemini") {
    GeminiCLIDriver$new(model = "gemini-1.5-flash")
  } else {
    NULL
  }

  AgentLLMNode$new(
    id = id,
    label = label,
    role = story_logic_registry$roles[[id]],
    driver = driver_obj,
    prompt_builder = story_logic_registry$prompts[[id]]
  )
}
```

## Building the DAG via Mermaid

We define the entire workflow architecture as a Mermaid string. This string demonstrates the **Fan-Out** from the `director` to the three writers, and the **Fan-In** from the writers to the `editor`.

```{r mermaid_source}
mermaid_graph <- "
graph TD
  director[Director | driver=gemini] --> writer_action[Action Writer | driver=gemini]
  director --> writer_mystery[Mystery Writer | driver=gemini]
  director --> writer_romance[Romance Writer | driver=gemini]

  writer_action --> editor[Master Editor | driver=gemini]
  writer_mystery --> editor
  writer_romance --> editor
"

# Instantiate the DAG
dag <- AgentDAG$from_mermaid(mermaid_graph, node_factory = story_node_factory)
compiled_dag <- dag$compile()
```

## Building and Running the DAG

We assemble the node into an `AgentDAG` and execute it.

```{r running}
# Execute with the initial premise from the registry
res <- compiled_dag$run(initial_state = story_logic_registry$initial_state)

# View the final synthesized story
cat(res$results$editor$output)
```

## Conclusion

The Fan-Out / Fan-In pattern is a robust way to generate diverse perspectives or sub-components of a task in parallel before synthesizing them. Using `HydraR`'s native DAG execution, these complex multi-agent flows are simple to define and orchestrate.

<!-- APAF Bioinformatics | fan_out_fan_in.Rmd | Approved | 2026-03-31 -->
