---
title: "Gemini CLI Driver Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gemini CLI Driver Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette demonstrates how to use the `GeminiCLIDriver` to power an agentic node in `HydraR`. 

## Setup

First, we define the `GeminiCLIDriver`. This driver interacts with the `gemini` command-line tool.

```{r setup}
library(HydraR)

# Initialize the Gemini CLI driver
driver <- GeminiCLIDriver$new(model = "gemini-1.5-flash")
```

## Defining the Agent Node

We create an `AgentLLMNode` that acts as a poetic bioinformatician.

```{r node}
node_agent <- AgentLLMNode$new(
  id = "writer",
  role = "You are a poetic assistant specializing in bioinformatics.",
  driver = driver,
  prompt_builder = function(state) {
    sprintf("Write a 2-line poem about: %s", state$get("topic"))
  }
)
```

## Building and Running the DAG

We assemble the node into an `AgentDAG` and execute it.

```{r running}
# Assemble the DAG
dag <- AgentDAG$new()
dag$add_node(node_agent)
dag$compile()

# Execute with an initial topic
res <- dag$run(initial_state = list(topic = "DNA sequencing"))

# View the output
cat(res$results$writer$output)
```

<!-- APAF Bioinformatics | gemini_cli_demo.Rmd | Approved | 2026-03-29 -->
