---
title: "Isolated Execution with Git Worktrees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Isolated Execution with Git Worktrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

A major challenge in building autonomous agents is **isolation**. When an agent modifies a file (e.g., writing a report or fixing a bug), it can easily cause conflicts with other parallel agents or the user's main working directory.

`HydraR` solves this by natively integrating with **Git Worktrees**. This guide demonstrates a "Toy Program" where an agent creates a new file in an isolated worktree.

---

## ­¤ÅŚ’ĖÅ Step 1: Initialize a Temporary Git Repository

To avoid modifying your actual project, we'll create a temporary Git repository in your `tempdir()`.

```{r setup, eval = FALSE}
library(HydraR)
library(withr)

# 1. Create a temporary folder
repo_root <- file.path(tempdir(), "hydra-toy-repo")
dir.create(repo_root)

# 2. Initialize Git and a README
withr::with_dir(repo_root, {
  system("git init -b main")
  system("git config user.email 'apaf@example.com'")
  system("git config user.name 'APAF Agent'")
  writeLines("# Toy Project", "README.md")
  system("git add README.md")
  system("git commit -m 'Initial commit'")
})
```

---

## ­¤ż¢ Step 2: Define a Writing Agent

We'll use the `GeminiCLIDriver` to request code generation. The agent will attempt to write a simple R function into a file called `math_fun.R`.

```{r define_dag, eval = FALSE}
# 1. Set up the Driver
# Requires 'gemini' CLI to be installed
driver <- GeminiCLIDriver$new(id = "gemini_bot")
register_driver(driver)

# 2. Define the Writer node
node_writer <- AgentLLMNode$new(
  id = "code_creator",
  driver_id = "gemini_bot",
  prompt_template = "Generate a simple R function to calculate the square of a number. SAVE IT to a file named 'math_fun.R'. Output ONLY the code."
)

# 3. Define the Merger node (Harmonizer)
# This node will reconcile the worktree changes back to 'main'
node_merger <- create_merge_harmonizer(id = "merger")

# 4. Build the DAG
dag <- AgentDAG$new()
dag$add_node(node_writer)
dag$add_node(node_merger)
dag$add_edge("code_creator", "merger")
dag$compile()
```

---

## ­¤øĪ’ĖÅ Step 3: Run with Worktree Isolation

By setting `use_worktrees = TRUE`, `HydraR` will:
1.  **Spawn a Worktree**: Create a new git branch and directory for the `code_creator` node.
2.  **Execute Isolated**: The Gemini CLI sees the worktree directory but **cannot** see your main repo files.
3.  **Harmonize**: The `merger` node will merge the `math_fun.R` file back to the `main` branch.

```{r run_dag, eval = FALSE}
# Run the DAG from the temporary repo root
results <- dag$run(
  initial_state = list(input = "Start execution"),
  use_worktrees = TRUE,
  repo_root = repo_root,
  fail_if_dirty = FALSE # Allow running in the new repo
)

# Verify the file was merged into the main repo branch
list.files(repo_root)
# [1] "README.md" "math_fun.R"
```

---

## ­¤¦ś Why Use Worktrees?

1.  **No Conflicts**: Two parallel agents can modify `DESCRIPTION` at the same time without corrupted git states.
2.  **Explicit Merging**: Changes only appear in your main directory when a `MergeHarmonizer` node explicitly merges them or the DAG succeeds.
3.  **Hardened Context**: If an agent runs `system("rm -rf *")`, it only destroys its temporary worktree, leaving your primary codebase safe.

---
<!-- APAF Bioinformatics | git_worktree_toy.Rmd | Approved | 2026-03-29 -->
