---
title: "Isolated Execution with Git Worktrees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Isolated Execution with Git Worktrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

A major challenge in building autonomous agents is **isolation**. When an agent modifies a file (e.g., writing a report or fixing a bug), it can easily cause conflicts with other parallel agents or the user's main working directory.

`HydraR` solves this by natively integrating with **Git Worktrees**. This guide demonstrates a "Toy Program" where an agent creates a new file in an isolated worktree.

---

## 🏗️ Step 1: Initialize a Temporary Git Repository

To avoid modifying your actual project, we'll create a temporary Git repository in your `tempdir()`.

```{r setup, eval = FALSE}
library(HydraR)
library(withr)

# 1. Create a temporary folder
repo_root <- file.path(tempdir(), "hydra-toy-repo")
dir.create(repo_root)

# 2. Initialize Git and a README
withr::with_dir(repo_root, {
  system("git init -b main")
  system("git config user.email 'apaf@example.com'")
  system("git config user.name 'APAF Agent'")
  writeLines("# Toy Project", "README.md")
  system("git add README.md")
  system("git commit -m 'Initial commit'")
})
```

---

## 🤖 Step 2: Define the Workflow Components

To keep our architecture clean, we store all workflow components—agent roles and prompts—in a central registry.

```{r logic_registry, eval = FALSE}
git_logic_registry <- list(
  # 1. Agent Roles
  roles = list(
    code_creator = "Generate a simple R function to calculate the square of a number. SAVE IT to a file named 'math_fun.R'. Output ONLY the code."
  ),

  # 2. Prompt Builders
  prompts = list(
    code_creator = function(state) {
      # In this example, we use the role as the direct prompt for simplicity
      git_logic_registry$roles$code_creator
    }
  )
)
```

## The Node Factory

We use a factory function to dynamically resolve nodes. This factory handles both standard LLM nodes and specialized nodes like the **Merge Harmonizer**.

```{r factory, eval = FALSE}
git_node_factory <- function(id, label, params) {
  if (id == "merger") {
    # Resolve the specialized MergeHarmonizer node
    create_merge_harmonizer(id = id)
  } else {
    # Resolve standard LLM nodes
    driver_obj <- if (!is.null(params$driver) && params$driver == "gemini") GeminiCLIDriver$new() else NULL

    AgentLLMNode$new(
      id = id,
      label = label,
      role = git_logic_registry$roles[[id]],
      driver = driver_obj,
      prompt_builder = git_logic_registry$prompts[[id]]
    )
  }
}
```

## Building the DAG via Mermaid

We define the entire workflow architecture as a Mermaid string. This string serves as the single source of truth for both structure and node metadata.

```{r mermaid_source, eval = FALSE}
mermaid_graph <- "
graph TD
  code_creator[Code Creator | driver=gemini] --> merger
  merger[Merge Harmonizer]
"

# Instantiate the DAG
dag <- AgentDAG$from_mermaid(mermaid_graph, node_factory = git_node_factory)
compiled_dag <- dag$compile()
```

---

## 🛡️ Step 3: Run with Worktree Isolation

By setting `use_worktrees = TRUE`, `HydraR` will:
1.  **Spawn a Worktree**: Create a new git branch and directory for the `code_creator` node.
2.  **Execute Isolated**: The Gemini CLI sees the worktree directory but **cannot** see your main repo files.
3.  **Harmonize**: The `merger` node will merge the `math_fun.R` file back to the `main` branch.

```{r run_dag, eval = FALSE}
# Run the DAG from the temporary repo root
results <- compiled_dag$run(
  initial_state = list(input = "Start execution"),
  use_worktrees = TRUE,
  repo_root = repo_root,
  fail_if_dirty = FALSE # Allow running in the new repo
)

# Verify the file was merged into the main repo branch
list.files(repo_root)
# [1] "README.md" "math_fun.R"
```

---

## 🧘 Why Use Worktrees?

1.  **No Conflicts**: Two parallel agents can modify `DESCRIPTION` at the same time without corrupted git states.
2.  **Explicit Merging**: Changes only appear in your main directory when a `MergeHarmonizer` node explicitly merges them or the DAG succeeds.
3.  **Hardened Context**: If an agent runs `system("rm -rf *")`, it only destroys its temporary worktree, leaving your primary codebase safe.

---
<!-- APAF Bioinformatics | git_worktree_toy.Rmd | Approved | 2026-03-29 -->
