---
title: "Hello World: Building an Agentic Loop with HydraR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello World: Building an Agentic Loop with HydraR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(HydraR)
```

## Introduction

This vignette demonstrates how to construct a simple "Hello World" agentic loop using the `HydraR` framework and the **Gemini CLI**.

In this example, we will create a Directed Acyclic Graph (DAG) with two nodes:
1. An **Executor** node (using the Gemini LLM).
2. A **Validator** node ensuring the Executor achieved its goal.

## 1. Setup

First, load the library and the Gemini CLI driver.

```{r library_load}
library(HydraR)
```

## Defining the Workflow Components

To keep our architecture clean, we store all workflow components—initial configuration, LLM prompts, agent roles, and deterministic logic—in a central registry.

```{r logic_registry}
hello_logic_registry <- list(
  # 0. Initial Configuration
  initial_state = list(),

  # 1. Deterministic Logic Functions
  logic = list(
    Validator = function(state, params) {
      guess <- tolower(trimws(state$get("Guesser")))
      is_valid <- (guess == "hello")
      list(status = "SUCCESS", output = list(valid = is_valid))
    }
  ),

  # 2. LLM Agent Roles
  roles = list(
    Guesser = "You are an agent trying to guess a specific word. Your task is to output exactly one word in lowercase."
  ),

  # 3. LLM Prompt Builders
  prompts = list(
    Guesser = function(state) {
      "Guess the secret greeting word."
    }
  )
)
```

## The Node Factory

We use a factory function to dynamically create nodes based on their type and parameters defined in the Mermaid graph.

```{r factory}
hello_node_factory <- function(id, label, params) {
  # Driver resolution from Mermaid params
  driver_obj <- if (!is.null(params$driver) && params$driver == "gemini") GeminiCLIDriver$new() else NULL

  if (id %in% names(hello_logic_registry$logic)) {
    # Create a deterministic Logic Node
    AgentLogicNode$new(
      id = id,
      label = label,
      logic_fn = hello_logic_registry$logic[[id]]
    )
  } else {
    # Create an agentic LLM Node
    AgentLLMNode$new(
      id = id,
      label = label,
      role = hello_logic_registry$roles[[id]],
      driver = driver_obj,
      prompt_builder = hello_logic_registry$prompts[[id]]
    )
  }
}
```

## Building the DAG via Mermaid

We define the entire workflow architecture as a Mermaid string. This string serves as the single source of truth for both structure and node metadata.

```{r mermaid_source}
mermaid_graph <- "
graph TD
  Guesser[LLM Guesser | driver=gemini] --> Validator
  Validator[Greeting Validator] -- Incorrect --> Guesser
"

# Instantiate the DAG
dag <- AgentDAG$from_mermaid(mermaid_graph, node_factory = hello_node_factory)

# Add conditional logic for the guess loop
dag$add_conditional_edge(
  from = "Validator",
  test = function(out) isTRUE(out$valid),
  if_true = NULL, # Success!
  if_false = "Guesser" # Incorrect, try again
)

compiled_dag <- dag$compile()
```

## 5. Visualizing the Workflow

```{r, results = 'asis'}
cat("```mermaid\n")
cat(dag$compile()$plot(type = "mermaid"))
cat("\n```\n")
```

## 6. Execution

Run the DAG!

```{r, eval = FALSE}
final_state <- compiled_dag$run(
  initial_state = hello_logic_registry$initial_state,
  max_steps = 10
)

# View final results
cat("Final Guess:", final_state$state$get("Guesser"), "\n")
cat("Success Status:", final_state$state$get("valid"), "\n")
```

The system successfully loops until the LLM produces the correct target greeting!

<!-- APAF Bioinformatics | hello_world_agent.Rmd | Approved | 2026-03-29 -->
