---
title: "Hello World: Building an Agentic Loop with HydraR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello World: Building an Agentic Loop with HydraR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates how to construct a simple "Hello World" agentic loop using the `HydraR` framework. 
In this example, we will create a Directed Acyclic Graph (DAG) with two nodes:
1. An **Executor** node (simulating an LLM or an external tool).
2. A **Validator** node ensuring the Executor achieved its goal.

## 1. Setup

First, load the library:

```{r}
library(HydraR)
```

## 2. Initialize the DAG

We initialize an `AgentDAG` which will orchestrate our nodes.

```{r}
dag <- AgentDAG$new()
```

## 3. Define Nodes

We can use `AgentNode` or subclass it (such as `AgentLogicNode` from downstream implementations) to perform custom behaviors.
Here, we simulate an agent trying to guess a specific word.

```{r}
# Simulated LLM Node
executor_node <- AgentLogicNode$new(id = "Guesser", logic_fn = function(state, memory = NULL) {
  attempts <- state$get("attempts")
  if (is.null(attempts)) attempts <- 0
  attempts <- attempts + 1

  # Simulate the agent guessing the right word on the 3rd attempt
  guess <- if (attempts >= 3) "hello" else "hi"

  list(
    status = "SUCCESS",
    output = list(guess = guess, attempts = attempts)
  )
})

# Simulated Validator Node
validator_node <- AgentLogicNode$new(id = "Validator", logic_fn = function(state, memory = NULL) {
  guess <- state$get("guess")
  is_valid <- (guess == "hello")
  list(
    status = "SUCCESS",
    output = list(valid = is_valid)
  )
})

dag$add_node(executor_node)
dag$add_node(validator_node)
dag$set_start_node("Guesser")
```

## 4. Define Transitions

We create a conditional transition to loop back if the guess is incorrect.

```{r}
# Unconditional edge from Guesser to Validator
dag$add_edge("Guesser", "Validator")

# Conditional edge: if not valid, loop back to Guesser
dag$add_conditional_edge("Validator",
  test = function(out) isTRUE(out$valid),
  if_true = NULL, # Stop execution on success
  if_false = "Guesser" # Loop back if not valid
)
```

## 5. Execution

Run the DAG!

```{r}
final_state <- dag$run(initial_state = list(attempts = 0), max_steps = 10)

# View final state
cat("Total Attempts:", final_state$state$get("attempts"), "\n")
cat("Final Guess:", final_state$state$get("guess"), "\n")
```

<!-- APAF Bioinformatics | hello_world_agent.Rmd | Approved | 2026-03-29 -->
