---
title: "Hello World Logic Nodes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello World Logic Nodes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

This vignette provides a basic "Hello World" example of `HydraR` logic nodes, demonstrating simple data processing and state persistence.

## Setup

```{r setup}
library(HydraR)
```

## Defining Logic Nodes

Logic nodes perform computation on the project state. We define two: one to collect input and another to transform it.

```{r nodes}
# 1. Collect Input Node
node_input <- AgentLogicNode$new(
  id = "collect_input",
  logic_fn = function(state) {
    list(status = "SUCCESS", output = list(input_raw = state$get("input")))
  }
)

# 2. Process Data Node
node_process <- AgentLogicNode$new(
  id = "process_data",
  logic_fn = function(state) {
    raw <- state$get("input_raw")
    res <- paste0("HYDRAR says: ", toupper(raw))
    list(status = "SUCCESS", output = list(processed_result = res))
  }
)
```

## Assembling the DAG

Next, we add these nodes to an `AgentDAG` and connect them.

```{r dag}
dag <- AgentDAG$new()
dag$add_node(node_input)
dag$add_node(node_process)
dag$add_edge("collect_input", "process_data")
dag$compile()
```

## Running with the Checkpointer

We use a `MemorySaver` to persist the state throughout the run.

```{r running}
checkpointer <- MemorySaver$new()
thread_id <- "hello_test_run"

final <- dag$run(
  initial_state = list(input = "hello hydra"),
  checkpointer = checkpointer,
  thread_id = thread_id
)

print(final$results$process_data$output$processed_result)
```

## Verifying Checkpoint Restoration

The `Checkpointer` allows us to verify the final state.

```{r restore}
restored <- checkpointer$get(thread_id)
print(restored$get("processed_result"))
```

<!-- APAF Bioinformatics | hello_world_logic.Rmd | Approved | 2026-03-29 -->
