---
title: "Agentic Travel Planning: Sydney to Hong Kong"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Travel Planning: Sydney to Hong Kong}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Zero-R-Code** orchestration pattern using `HydraR`. Instead of defining roles, logic, and state in R blocks, we define the entire workflow in a single `workflow.yml` file.

The R environment acts purely as an **interpreter/compiler** for the language-independent manifest.

## Setup

First, load the HydraR library.

```{r lib_load}
library(HydraR)
```

## Loading the Workflow (Declarative YAML)

We use `load_workflow()` to ingest the entire definition from `hong_kong_travel.yml`. This file contains the:
1. **Mermaid Graph** (Source of truth for DAG architecture)
2. **LLM Roles** (System prompts for travel concierges)
3. **Anonymous Logic** (R snippets for constraints and prompting)
4. **Initial State** (Pre-configurations for the journey)

```{r load_wf}
# Load everything from the external declarative source
wf <- load_workflow("hong_kong_travel.yml")
```

## Instantiating the DAG

With the registries populated by the loader, we simply pass the graph and the universal `auto_node_factory()` to `mermaid_to_dag()`.

```{r dag_creation}
dag <- spawn_dag(wf, auto_node_factory())
```

## Visualizing the Workflow

We can view the agent's logic directly using Mermaid.js syntax.

```{r, results = 'asis'}
cat("```mermaid\n")
cat(dag$plot(type = "mermaid", details = TRUE))
cat("\n```\n")
```

## Execution

When we run the DAG, we use the `initial_state` extracted from the YAML file. No manual R list creation is required.

```{r eval=TRUE}
# Register a checkpointer for durability
checkpointer <- DuckDBSaver$new(db_path = "travel_booking.duckdb")

# Toggle for slow/expensive and potentially rate-limited AI figure generation
# Set to TRUE to force-regenerate the images, otherwise uses existing ones.
FORCE_REGENERATE_IMAGES <- FALSE
ASPECT_RATIO <- "16:9"

# Run the orchestration using the state from YAML
results <- dag$run(
  initial_state = append(wf$initial_state, list(
    force_regenerate_images = FORCE_REGENERATE_IMAGES,
    aspect_ratio = ASPECT_RATIO
  )),
  max_steps = 15,
  checkpointer = checkpointer
)

# Display final itinerary
cat("\n\n### Generated Itinerary\n")
cat(as.character(results$state$get("Planner")))

# Display Constraint Audit Report
cat("\n\n### Constraint Audit Report\n")
report <- results$state$get("report")
if (!is.null(report)) {
  cat(as.character(report))
} else {
  cat("No audit report available.")
}

# Display Pamphlet (HTML)
cat("\n\n### Formatted Pamphlet (HTML Fragment)\n")
pamphlet_html <- results$state$get("PamphletFormatter")
if (!is.null(pamphlet_html)) {
  htmltools::HTML(pamphlet_html)
} else {
  cat("Pamphlet not generated.")
}

# List Artifacts
cat("\n\n### Generated Artifacts\n")
artifacts <- list.files(pattern = "hong_kong|validation_report")
if (length(artifacts) > 0) {
  cat(paste("- ", artifacts, collapse = "\n"))
} else {
  cat("No artifacts found.")
}
```

## Conclusion

This workflow demonstrates how `HydraR` enables **Truly Zero-R-Code** definitions:
1. **Language-Independent**: The workflow is defined in YAML and Mermaid, making it portable.
2. **Reduced Boilerplate**: `load_workflow()` handles all registration and state parsing.
3. **Maintainable**: Logic and roles are separated from the R execution engine.

---
<!-- APAF Bioinformatics | hong_kong_travel.Rmd | Approved | 2026-03-30 -->
