---
title: "Agentic Travel Planning: Sydney to Hong Kong"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Travel Planning: Sydney to Hong Kong}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

## Introduction

This vignette demonstrates how to use `HydraR` to orchestrate a high-fidelity travel planning workflow. Unlike simple mock-based agents, we will use the `GeminiCLIDriver` to drive real-world logic transitions, state management, and persistence for a complex itinerary.

We will plan a trip from **Sydney (SYD)** to **Hong Kong (HKG)** with specific airline and activity constraints.

## Scenario Requirements

- **Route**: Sydney to Hong Kong (Departure: 29th May 2026, Return: 5th June 2026).
- **Airline**: Qantas (Required).
- **Activities**: Visit Cheung Chau Island, dine at Spaghetti House, and sample local Cantonese cuisine.
- **Goal**: Generate a structured 3-day itinerary that respects these constraints.

## Setup

First, load the library and initialize the `GeminiCLIDriver`.

```{r}
library(HydraR)

# Initialize the Gemini CLI driver
# Note: This assumes the 'gemini' CLI is installed and configured on your system.
driver <- GeminiCLIDriver$new()
```

## Defining the Agent State

We define the initial state, which includes our hard constraints and preferences.

```{r}
initial_state <- list(
    origin = "Sydney",
    destination = "Hong Kong",
    departure_date = "2026-05-26",
    return_date = "2026-06-01",
    airline = "Qantas",
    must_include = c("Cheung Chau Island", "Spaghetti House", "Local Cuisine"),
    itinerary_draft = NULL,
    validation_passed = FALSE
)

state <- AgentState$new(initial_state)
```

## Building the Orchestration Graph

We will define two nodes:
1. **The Planner**: An `AgentLLMNode` that calls Gemini to generate the itinerary.
2. **The Auditor**: An `AgentLogicNode` that checks if all "must-include" items are actually in the draft.

### 1. The Planner Node

We use a `prompt_builder` to inject the state dynamically into the LLM prompt.

```{r}
planner_node <- AgentLLMNode$new(
    id = "TravelPlanner",
    label = "Gemini Itinerary Engine",
    role = "You are a professional travel concierge specializing in premium Hong Kong travel.",
    driver = driver,
    prompt_builder = function(state) {
        sprintf(
            "Plan a trip from %s to %s.
            Dates: %s to %s.
            Airline: %s.
            Must include: %s.
            Provide a 3-day detailed itinerary in Hong Kong.",
            state$get("origin"),
            state$get("destination"),
            state$get("departure_date"),
            state$get("return_date"),
            state$get("airline"),
            paste(state$get("must_include"), collapse = ", ")
        )
    }
)
```

### 2. The Auditor Node

This node uses standard R logic to validate the LLM's output.

```{r}
auditor_node <- AgentLogicNode$new(
    id = "Auditor",
    label = "Constraint Validator",
    logic_fn = function(state) {
        itinerary <- state$get("output") # From the previous LLM node
        must_include <- state$get("must_include")

        # Simple string matching to check constraints
        found <- sapply(must_include, function(x) grepl(x, itinerary, ignore.case = TRUE))

        if (all(found)) {
            list(
                status = "SUCCESS",
                output = list(validation_passed = TRUE, message = "All constraints met!")
            )
        } else {
            missing <- must_include[!found]
            list(
                status = "SUCCESS",
                output = list(
                    validation_passed = FALSE,
                    message = paste("Missing items:", paste(missing, collapse = ", "))
                )
            )
        }
    }
)
```

## Compiling the DAG

We link the nodes and add a conditional loop. If the Auditor fails, we go back to the Planner with the missing feedback.

```{r}
dag <- AgentDAG$new()
dag$add_node(planner_node)
dag$add_node(auditor_node)

dag$add_edge("TravelPlanner", "Auditor")

# Conditional loop: Back to planner if validation fails
dag$add_conditional_edge(
    from = "Auditor",
    test = function(out) isTRUE(out$validation_passed),
    if_true = NULL, # END
    if_false = "TravelPlanner"
)

dag$set_start_node("TravelPlanner")
compiled_dag <- dag$compile()
```

## Visualizing the Workflow

We can view the agent's logic directly using Mermaid.js syntax.

```{r, results = 'asis'}
cat("```mermaid\n")
cat(compiled_dag$plot(type = "mermaid"))
cat("\n```\n")
```

## Execution

When we run the DAG, `HydraR` manages the state transitions and LLM calls. If the LLM forgets to mention "Spaghetti House", the Auditor will detect it and loop back for a correction.

```{r, eval = FALSE}
# Register a checkpointer for durability
checkpointer <- AgentCheckpointer$new(path = "travel_booking.duckdb")
compiled_dag$set_checkpointer(checkpointer)

# Run the orchestration
results <- compiled_dag$run(initial_state = initial_state)

# Display final itinerary
cat(results$state$get("output"))
```

## Conclusion

This workflow demonstrates how `HydraR` provides:
1. **Dynamic Prompting**: Injecting complex R state into LLM calls.
2. **Reliability**: Using logic nodes as "guardrails" for non-deterministic LLM output.
3. **Persistence**: The ability to checkpoint and resume complex multi-step planning.

---
<!-- APAF Bioinformatics | hong_kong_travel.Rmd | Approved | 2026-03-29 -->
