---
title: "Agentic Personalized Shopping Assistant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Personalized Shopping Assistant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Personalized Shopping Assistant** pattern using `HydraR` and the **Gemini CLI**.

We define an agentic graph where a recommendation engine interacts with a mock user. 

The two roles are:
1. **Shopper Node**: Acts as the intelligent proxy that searches the catalogue based on preferences.
2. **User Proxy Node**: Represents the user feedback loop. It provides critiques on the recommended item.

This models a continuous interaction loop where the agent refines its search based on real-time feedback.

## Setup

```{r}
library(HydraR)

# Initialize the Gemini CLI driver
driver <- GeminiCLIDriver$new()
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Shopper Node

This node uses user feedback to update its recommendations.

```{r}
shopper_node <- AgentLLMNode$new(
  id = "Shopper",
  label = "Store Concierge",
  role = "You are a personalized shopping assistant. Based on the user's initial request and any feedback provide a specific product recommendation.",
  driver = driver,
  prompt_builder = function(state) {
    feedback_text <- if (!is.null(state$get("UserProxy"))) sprintf("\nUser Feedback: %s", state$get("UserProxy")) else ""
    sprintf("Original Request: %s%s\nOutput exactly a product name.", state$get("shopping_request"), feedback_text)
  }
)

dag$add_node(shopper_node)
```

### 2. The User Proxy Node

This node simulates a picky user receiving recommendations.

```{r}
user_proxy_node <- AgentLLMNode$new(
  id = "UserProxy",
  label = "Customer Proxy",
  role = "You are a customer looking for a specific item. Review the recommended product and either say 'I'll buy it' or provide 'Modification Feedback'.",
  driver = driver,
  prompt_builder = function(state) {
    sprintf("Recommended Item: %s\nOriginal Request: %s\nDecide if you like it or need something different.", state$get("Shopper"), state$get("shopping_request"))
  }
)

dag$add_node(user_proxy_node)
```

## Defining Transitions

Configure the feedback loop.

```{r}
dag$set_start_node("Shopper")

dag$add_edge("Shopper", "UserProxy")

dag$add_conditional_edge(
  from = "UserProxy",
  test = function(out) {
    grepl("buy it", out, ignore.case = TRUE)
  },
  if_true = NULL, # Done
  if_false = "Shopper" # Try again
)

compiled_dag <- dag$compile()
```

## Visualizing the Workflow

```{r, results = 'asis'}
cat("```mermaid\n")
cat(compiled_dag$plot(type = "mermaid"))
cat("\n```\n")
```

## Running the Scenario

Provide the initial request and observe the iterations.

```{r, eval = FALSE}
initial_state <- list(
  shopping_request = "I need a cool t-shirt for the weekend with graphic design."
)

cat("Starting Personalized Shopper...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 10)

cat("\n--- SHOPPING RESULT ---\n")
cat("Final Recommendation:", result$state$get("Shopper"), "\n")
cat("User Feedback:", result$state$get("UserProxy"), "\n")
```

The DAG successfully refined its recommendation through the interactive cycles until the customer proxy was satisfied!

<!-- APAF Bioinformatics | personalized_shopping.Rmd | Approved | 2026-03-29 -->
