---
title: "Agentic Software Bug Assistant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Software Bug Assistant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Software Bug Assistant** pattern using `HydraR`. 
In this scenario, two agent nodes work together to resolve a software bug:
1. **Bug Analyzer Node**: Reviews the error message and current code, then proposes a patch.
2. **Test Runner Node**: Simulates running a test suite against the proposed patch. If the test fails, it provides error traces back to the Bug Analyzer.

This is a classic "Generate-and-Test" self-healing loop.

## Setup

```{r}
library(HydraR)
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Bug Analyzer Node

This node mocks an LLM digesting a bug report and iterative test failures to eventually produce a working patch.

```{r}
analyzer_node <- AgentLogicNode$new(id = "Analyzer", logic_fn = function(state, memory = NULL) {
    # Extract state
    bug_report <- state$get("bug_report")
    test_feedback <- state$get("test_feedback")
    
    attempts <- state$get("patch_attempts")
    if (is.null(attempts)) attempts <- 0
    attempts <- attempts + 1
    
    # Mock behavior: It takes 2 tries to fix a Null Pointer Exception
    proposed_patch <- if (attempts == 1) {
        "if (x == NULL) return 0;" # Close, but maybe not quite right
    } else {
        "if (is.null(x)) return(0)" # Correct R syntax
    }
    
    list(
        status = "SUCCESS",
        output = list(
            proposed_patch = proposed_patch,
            patch_attempts = attempts
        )
    )
})

dag$add_node(analyzer_node)
```

### 2. The Test Runner Node

This node evaluates the patch. We'll simply check if the proposed patch is exactly the "correct R syntax" string.

```{r}
tester_node <- AgentLogicNode$new(id = "Tester", logic_fn = function(state, memory = NULL) {
    patch <- state$get("proposed_patch")
    
    if (patch == "if (is.null(x)) return(0)") {
        list(
            status = "SUCCESS",
            output = list(
                tests_passed = TRUE,
                test_feedback = "All 5 tests passed successfully."
            )
        )
    } else {
        list(
            status = "SUCCESS",
            output = list(
                tests_passed = FALSE,
                test_feedback = "Error: object 'NULL' not found. Did you mean to use 'is.null()'?"
            )
        )
    }
})

dag$add_node(tester_node)
```

## Defining Transitions

We structure the loop. `Tester` conditionally loops back to `Analyzer` upon test failure.

```{r}
dag$set_start_node("Analyzer")

dag$add_edge("Analyzer", "Tester")

dag$add_conditional_edge(
    from = "Tester",
    test = function(out) {
        isTRUE(out$tests_passed)
    },
    if_true = NULL,         # End execution
    if_false = "Analyzer"   # Retry
)

compiled_dag <- dag$compile()
```

## Running the Scenario

Provide the initial bug report and start the DAG.

```{r}
initial_state <- list(
    bug_report = "Function crashes when 'x' is missing."
)

cat("Starting Automatic Bug Remediation...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 10)

cat("\n--- RESOLUTION RESULT ---\n")
cat("Patch Attempts:", result$state$get("patch_attempts"), "\n")
cat("Final Patch:", result$state$get("proposed_patch"), "\n")
cat("Test Results:", result$state$get("test_feedback"), "\n")
```

The DAG correctly detected the test failure, routed the detailed feedback back to the analyzer, and produced a successful patch on its second attempt!

<!-- APAF Bioinformatics | software_bug_assistant.Rmd | Approved | 2026-03-29 -->
