---
title: "Parallel Benchmarking with Git Worktrees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel Benchmarking with Git Worktrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

Modern agentic workflows often involve multiple specialized agents working on different parts of a codebase simultaneously. When these agents modify files, they risk stepping on each other's toes—causing merge conflicts or corrupted git states.

`HydraR` addresses this by leveraging **Git Worktrees** for isolated execution. This guide demonstrates a parallel workflow where:
1.  **Three Agents** simultaneously write R code for different sorting algorithms (Bubble, Quick, Merge sort) in separate worktrees.
2.  A **Merge Harmonizer** automatically merges these independent branches back into the main repository.
3.  A **Benchmarking Node** executes the generated code to compare their performance.
4.  A **Visualization Node** plots the results.

---

## 🏗️ Step 1: Initialize an Isolated Repository

To demonstrate the power of worktrees, we first create a temporary Git repository.

```{r setup, eval = FALSE}
library(HydraR)
library(withr)
library(ggplot2)
library(future)

# 0. Setup Parallel Execution (for Worktree Isolation)
plan(multisession, workers = 3)
options(future.rng.onMisuse = "ignore")

# 1. Create a temporary folder
repo_root <- file.path(tempdir(), "sorting-benchmark-repo")
if (dir.exists(repo_root)) unlink(repo_root, recursive = TRUE)
dir.create(repo_root)

# 2. Initialize Git and a README
withr::with_dir(repo_root, {
  system("git init -b main")
  system("git config user.email 'apaf@example.com'")
  system("git config user.name 'APAF Agent'")
  system("git config commit.gpgsign false")
  writeLines("# Sorting Benchmark", "README.md")
  system("git add README.md")
  system("git commit -m 'Initial commit'")
})
```

---

## 🧠 Step 2: Define Logic and Architecture

We decouple our workflow logic into a central registry. This keeps our prompt engineering and implementation details separate from the DAG architecture.

### The Logic Registry

```{r logic_registry, eval = FALSE}
sorting_registry <- list(
  # 1. Agent Roles
  roles = list(
    bubble = "Write an R function named `bubble_sort` that takes a numeric vector and returns it sorted using the bubble sort algorithm. Output ONLY the code, no markdown blocks.",
    quick = "Write an R function named `quick_sort` that takes a numeric vector and returns it sorted using the quick sort algorithm. Output ONLY the code, no markdown blocks.",
    merge = "Write an R function named `merge_sort` that takes a numeric vector and returns it sorted using the merge sort algorithm. Output ONLY the code, no markdown blocks."
  ),

  # 2. Mock Fallbacks (for when LLM is unavailable)
  mocks = list(
    bubble = "bubble_sort <- function(x) {
      n <- length(x)
      for (i in 1:(n - 1)) {
        for (j in 1:(n - i)) {
          if (x[j] > x[j + 1]) {
            tmp <- x[j]
            x[j] <- x[j + 1]
            x[j + 1] <- tmp
          }
        }
      }
      return(x)
    }",
    quick = "quick_sort <- function(x) {
      if (length(x) <= 1) return(x)
      pivot <- x[1]
      rest <- x[-1]
      sv <- rest[rest < pivot]
      gv <- rest[rest >= pivot]
      return(c(quick_sort(sv), pivot, quick_sort(gv)))
    }",
    merge = "merge_sort <- function(x) {
      if (length(x) <= 1) return(x)
      m <- length(x) %/% 2
      l <- merge_sort(x[1:m])
      r <- merge_sort(x[(m+1):length(x)])
      res <- numeric(length(x))
      li <- 1; ri <- 1; i <- 1
      while(li <= length(l) && ri <= length(r)) {
        if(l[li] <= r[ri]) { res[i] <- l[li]; li <- li + 1 }
        else { res[i] <- r[ri]; ri <- ri + 1 }
        i <- i + 1
      }
      if(li <= length(l)) res[i:length(x)] <- l[li:length(l)]
      if(ri <= length(r)) res[i:length(x)] <- r[ri:length(r)]
      return(res)
    }"
  )
)
```

### The Node Factory

The factory dynamically resolves Mermaid nodes into `HydraR` node objects. It handles the "Failsafe" logic by falling back to mocks if the LLM driver is unavailable.

```{r factory, eval = FALSE}
sorting_node_factory <- function(id, label, params) {
  # 1. Merge Harmonization
  if (id == "merger") {
    return(create_merge_harmonizer(id = id))
  }

  # 2. Benchmarking Node
  if (id == "benchmark") {
    return(AgentLogicNode$new(id, function(state) {
      # Ensure worktree is synced with the merged main branch
      system("git checkout main", ignore.stdout = TRUE, ignore.stderr = TRUE)

      # Source all generated files
      files <- list.files(pattern = "_sort.R")
      message(sprintf("[Benchmark] Found %d algorithm files: %s", length(files), paste(files, collapse = ", ")))
      purrr::walk(files, source)

      # Benchmark parameters
      n_elements <- 1000
      n_trials <- 5
      methods <- c("bubble", "quick", "merge")

      results_df <- purrr::map_df(methods, function(m) {
        func_name <- paste0(m, "_sort")
        if (!exists(func_name)) {
          return(NULL)
        }
        func <- get(func_name)

        times <- purrr::map_dbl(1:n_trials, function(i) {
          test_data <- rnorm(n_elements)
          start <- Sys.time()
          func(test_data)
          as.numeric(difftime(Sys.time(), start, units = "secs"))
        })

        data.frame(method = m, time = times)
      })

      list(status = "success", output = results_df)
    }))
  }

  # 3. Visualization Node
  if (id == "plot") {
    return(AgentLogicNode$new(id, function(state) {
      df <- state$get("benchmark")
      if (is.null(df) || nrow(df) == 0) {
        return(list(status = "failed", output = NULL, error = "No benchmark data found."))
      }

      # Suppress graphics device warnings by opening a null device
      # This prevents future-related warnings in parallel workers
      pdf(NULL)
      on.exit(if (dev.cur() > 1) dev.off())

      # Use .data pronoun to avoid 'no visible binding' warnings
      p <- ggplot(df, aes(x = .data$method, y = .data$time, fill = .data$method)) +
        geom_boxplot() +
        theme_minimal() +
        labs(title = "Sorting Algorithm Performance", y = "Time (seconds)", x = "Algorithm")

      # In a real vignette, we'd return the plot object
      # But here we show it.
      pdf(file = "sorting_benchmark.pdf")
      print(p)
      dev.off()
      list(status = "success", output = "Plot rendered.")
    }))
  }

  # 4. Sorting Agents (Parallel) using hardened AgentLLMNode
  # The parameters are parsed from the Mermaid string below
  AgentLLMNode$new(id,
    role = sorting_registry$roles[[id]],
    driver = GeminiCLIDriver$new(model = "gemini-2.5-flash"),
    label = label,
    params = params
  )
}
```

---

## 🤖 Step 3: Orchestration and Execution

We define our parallel DAG using Mermaid syntax. By connecting multiple agents to the `merger` node, `HydraR` knows to execute them in isolated worktrees.

```{r run_dag, eval = FALSE}
# 1. Define the Mermaid Architecture
# output_format=r triggers automatic code extraction in AgentLLMNode
# output_path ensures the extracted code is persisted to the filesystem
mermaid_graph <- '
graph TD
    bubble["Bubble Agent | output_format=r | output_path=bubble_sort.R"] --> merger
    quick["Quick Agent | output_format=r | output_path=quick_sort.R"] --> merger
    merge["Merge Agent | output_format=r | output_path=merge_sort.R"] --> merger
    merger["Merge Harmonizer"] --> benchmark
    benchmark["Benchmark Performance"] --> plot
    plot["Visualize Results"]
'

# 2. Instantiate and Compile
dag <- AgentDAG$from_mermaid(mermaid_graph, node_factory = sorting_node_factory)
compiled_dag <- dag$compile()

# 3. Run with Worktree Isolation
# use_worktrees = TRUE activates the WorktreeManager
results <- compiled_dag$run(
  initial_state = list(input = "Start sorting benchmark", repo_root = repo_root),
  use_worktrees = TRUE,
  repo_root = repo_root,
  fail_if_dirty = FALSE,
  packages = c("withr", "HydraR", "ggplot2")
)

# 4. Save Execution Trace
compiled_dag$save_trace("sorting_trace.json")
```

---

## 🧘 Summary

By using **Git Worktrees**, we successfully:
1.  **Eliminated State Corruption**: Three agents modified the codebase simultaneously without interfering with each other.
2.  **Automated Conflict Resolution**: The `MergeHarmonizer` handled the integration of disparate branches into the main codebase.
3.  **End-to-End Validation**: We used the generated code immediately in a downstream benchmarking task, closing the loop between "Agent Action" and "Scientific Result."

---
<!-- APAF Bioinformatics | sorting_benchmark.Rmd | Approved | 2026-03-30 -->
