---
title: "State Persistence & Restart: Resilient Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{State Persistence & Restart: Resilient Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(HydraR)
```

This vignette demonstrates a core feature of the `HydraR` framework: **Automated State Persistence & Restart**. In high-complexity multi-agent workflows, code can often fail due to environmental factors, API timeouts, or logic errors. `HydraR` addresses these issues by allowing you to checkpoint every step and resume exactly from where you left off.

## Persistence Concepts

1.  **Checkpointing**: After each node resolves, its entire state and output are saved to a persistent backend.
2.  **Pausing**: A node can return a `PAUSE` status, signaling the orchestrator to save the thread and wait for external feedback.
3.  **Thread Resumption**: Using a unique `thread_id`, you can reload a saved process and restart from any node.

## Declarative Pipeline

We define a 3-step pipeline with a "Risky" node in the middle that will pause until a condition is met.

```{r load, eval = TRUE}
library(HydraR)

# Load the resilient workflow from YAML
wf <- load_workflow("state_persistence.yml")

# Spawn and compile the DAG
dag <- spawn_dag(wf)
```

## Running the Scenario

### 1. Initial Run (The Pause)

We run the DAG with `fixed = FALSE`. The second node will **pause** the pipeline and save the state to DuckDB.

```{r run_1, eval = TRUE}
# Configure DuckDB Persistence
saver <- DuckDBSaver$new(db_path = "history.duckdb")
tid <- "session-001"

# Run 1: Expected to pause at Step2
res1 <- dag$run(
  thread_id = tid,
  checkpointer = saver,
  initial_state = list(fixed = FALSE)
)

print(res1$status) # "PAUSE"
```

### 2. The Restart (The Success)

Next, we "fix" the state by setting `fixed = TRUE` and restart from `Step2`. `HydraR` detects the checkpoint, restores the state from DuckDB, and re-executes from the paused node.

```{r run_2, eval = TRUE}
# Run 2: Resume from Step2 using the SAME thread_id
final_results <- dag$run(
  thread_id = tid,
  checkpointer = saver,
  initial_state = list(fixed = TRUE),
  resume_from = "Step2"
)

# Step 1 was skipped; the process started from Step 2!
print(final_results$status) # "SUCCESS"
```

## Summary

The `DuckDB` checkpointer ensures that no computational progress is lost. This is critical for agentic workflows involving expensive LLM calls or long-running bioinformatics pipelines where re-running the entire system from scratch is prohibitive.

---
<!-- APAF Bioinformatics | state_persistence.Rmd | Approved | 2026-04-03 -->
