---
title: "Agentic Story Teller"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Story Teller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Story Teller** pattern using `HydraR` and the **Gemini CLI**.

In this workflow, two LLM agents collaborate to write a story:
1. **Writer Agent**: Drafts the story based on the prompt or incorporates feedback.
2. **Reviewer Agent**: Critiques the draft using a rubric. If the story meets the criteria, it approves the draft. Otherwise, it sends feedback back to the Writer.

This forms a fundamental **iterative critique-and-revise** loop.

## Setup

```{r}
library(HydraR)

# Initialize the Gemini CLI driver
driver <- GeminiCLIDriver$new()
```

## Building the Story Teller DAG

We'll define an `AgentDAG` and inject two `AgentLLMNode` nodes to demonstrate the collaboration.

```{r}
dag <- AgentDAG$new()
```

### 1. The Writer Node

The `Writer` node drafts or improves the story based on reviewer feedback.

```{r}
writer_node <- AgentLLMNode$new(
  id = "Writer",
  label = "Story Author",
  role = "You are a creative writer. Draft a story based on the initial prompt or update it according to reviewer feedback. Output exactly the story draft.",
  driver = driver,
  prompt_builder = function(state) {
    feedback_text <- if (!is.null(state$get("Reviewer"))) sprintf("\nFeedback: %s", state$get("Reviewer")) else ""
    sprintf("Prompt: %s%s\nDraft the story.", state$get("story_prompt"), feedback_text)
  }
)

dag$add_node(writer_node)
```

### 2. The Reviewer Node

The `Reviewer` node checks the `story_draft`. It provides feedback or approves it.

```{r}
reviewer_node <- AgentLLMNode$new(
  id = "Reviewer",
  label = "Literary Editor",
  role = "You are a literary editor. Critically review the story draft for tone and quality. If it is excellent, say 'Approved'. Otherwise, give specific critique.",
  driver = driver,
  prompt_builder = function(state) {
    sprintf("Draft: %s\nReview this draft.", state$get("Writer"))
  }
)

dag$add_node(reviewer_node)
```

## Defining Transitions

We configure the graph to start with the Writer, unconditionally pass to the Reviewer, and conditionally loop back or finish.

```{r}
dag$set_start_node("Writer")

# Unconditional transition from Writer -> Reviewer
dag$add_edge("Writer", "Reviewer")

# Conditional transition from Reviewer -> Writer or END
dag$add_conditional_edge(
  from = "Reviewer",
  test = function(out) {
    grepl("Approved", out, ignore.case = TRUE)
  },
  if_true = NULL, # Stop execution
  if_false = "Writer" # Loop back
)
```

## Visualizing the Workflow

```{r, results = 'asis'}
cat("```mermaid\n")
cat(dag$compile()$plot(type = "mermaid"))
cat("\n```\n")
```

## Running the Agent

```{r, eval = FALSE}
compiled_dag <- dag$compile()

cat("Starting Collaborative Writing Process...\n")
result <- compiled_dag$run(initial_state = list(story_prompt = "A story about a robot learning to cook."), max_steps = 10)

cat("--- FINAL RESULT ---\n")
cat("Final Feedback:", result$state$get("Reviewer"), "\n")
cat("Final Story Draft:\n", result$state$get("Writer"), "\n")
```

As we can see, HydraR flawlessly orchestrated the stateful communication and iteration loop between the two agents to refine the creative output!

<!-- APAF Bioinformatics | story_teller.Rmd | Approved | 2026-03-29 -->
