---
title: "Agentic Story Teller"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Story Teller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates the **Story Teller** pattern using `HydraR`. 
In this workflow, two LLM agents collaborate to write a story:
1. **Writer Agent**: Drafts the story based on the prompt or incorporates feedback.
2. **Reviewer Agent**: Critiques the draft using a rubric. If the story meets the criteria, it approves the draft. Otherwise, it sends feedback back to the Writer.

This forms a fundamental **iterative critique-and-revise** loop.

## Setup

```{r}
library(HydraR)
```

## Building the Story Teller DAG

We'll define an `AgentDAG` and inject two custom nodes to mock the LLM behavior. In a real-world scenario, you would use `node_llm()` or `AgentLogicNode` bound to a specific LLM driver (e.g., Gemini).

```{r}
dag <- AgentDAG$new()
```

### 1. The Writer Node

The `Writer` node looks at the current `feedback` (if any) and updates the `story_draft`. We'll mock the behavior to show that the story gets slightly better over time.

```{r}
writer_node <- AgentLogicNode$new(id = "Writer", logic_fn = function(state, memory = NULL) {
  # Extract current state
  current_draft <- state$get("story_draft")
  iteration <- state$get("writer_iterations")
  if (is.null(iteration)) iteration <- 0
  feedback <- state$get("reviewer_feedback")

  # Increment iteration
  iteration <- iteration + 1

  # Generate new draft
  if (is.null(current_draft)) {
    new_draft <- "Once upon a time, a brave knight went on a quest."
    action_msg <- "Drafting initial story."
  } else {
    new_draft <- paste(current_draft, "He fought a dragon and saved the kingdom.")
    action_msg <- paste("Improving story based on feedback:", feedback)
  }

  list(
    status = "SUCCESS",
    output = list(
      story_draft = new_draft,
      writer_iterations = iteration,
      latest_action = action_msg
    )
  )
})

dag$add_node(writer_node)
```

### 2. The Reviewer Node

The `Reviewer` node checks the `story_draft`. In our mock logic, the reviewer demands at least 2 iterations before it considers the story complete.

```{r}
reviewer_node <- AgentLogicNode$new(id = "Reviewer", logic_fn = function(state, memory = NULL) {
  iteration <- state$get("writer_iterations")

  if (iteration >= 2) {
    # Story is approved!
    list(
      status = "SUCCESS",
      output = list(
        is_approved = TRUE,
        reviewer_feedback = "This is a great story! Approved."
      )
    )
  } else {
    # Needs more work
    list(
      status = "SUCCESS",
      output = list(
        is_approved = FALSE,
        reviewer_feedback = "The story lacks action. Add a dragon."
      )
    )
  }
})

dag$add_node(reviewer_node)
```

## Defining Transitions

We configure the graph to start with the Writer, unconditionally pass to the Reviewer, and conditionally loop back or finish.

```{r}
dag$set_start_node("Writer")

# Unconditional transition from Writer -> Reviewer
dag$add_edge("Writer", "Reviewer")

# Conditional transition from Reviewer -> Writer or END
dag$add_conditional_edge(
  from = "Reviewer",
  test = function(out) {
    # Return the boolean approval status
    isTRUE(out$is_approved)
  },
  if_true = NULL, # Stop execution
  if_false = "Writer" # Loop back to Writer
)
```

## Running the Agent

We compile and run the DAG with an empty initial state. We set a `max_steps` to guarantee the loop terminates securely (a built-in HydraR safeguard).

```{r}
compiled_dag <- dag$compile()

cat("Starting Collaborative Writing Process...\n")
result <- compiled_dag$run(initial_state = list(), max_steps = 10)

# Observe the final output
final_story <- result$state$get("story_draft")
total_iterations <- result$state$get("writer_iterations")
feedback <- result$state$get("reviewer_feedback")

cat("--- FINAL RESULT ---\n")
cat("Total Writer Iterations:", total_iterations, "\n")
cat("Final Feedback:", feedback, "\n")
cat("Final Story Draft:\n", final_story, "\n")
```

As we can see, HydraR flawlessly orchestrated the stateful communication and iteration loop between the two mock agents.

<!-- APAF Bioinformatics | story_teller.Rmd | Approved | 2026-03-29 -->
