---
title: "Agentic Travel Booking Concierge"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Agentic Travel Booking Concierge}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates how to orchestrate a **Travel Booking Concierge** workflow using `HydraR`.
A typical travel planning AI must satisfy various constraints such as the user's destination, duration, and crucially, their budget.

In this workflow, we define two nodes:
1. **Travel Planner Node**: Generates a proposed itinerary and an estimated cost based on user constraints and previous feedback.
2. **Itinerary Validator Node**: Receives the proposed itinerary, checks it against the user's hard budget constraints, and either approves it or demands a cheaper alternative.

This demonstrates robust condition handling through cyclic loops.

## Setup

```{r}
library(HydraR)
```

## Building the DAG

Initialize the `AgentDAG`.

```{r}
dag <- AgentDAG$new()
```

### 1. The Travel Planner Node

This node mocks an LLM calling flight and hotel APIs, receiving feedback if its previous proposal was too expensive.

```{r}
planner_node <- AgentLogicNode$new(id = "Planner", logic_fn = function(state, memory = NULL) {
  # Read user constraints
  destination <- state$get("destination")
  budget <- state$get("budget")

  # Check if we are iterating due to failure
  attempts <- state$get("planner_attempts")
  if (is.null(attempts)) attempts <- 0
  attempts <- attempts + 1

  # Mocking cost reduction on each attempt
  # Attempt 1: $2500, Attempt 2: $1800, Attempt 3: $1200
  base_cost <- 3200
  proposed_cost <- base_cost - (attempts * 700)

  itinerary <- sprintf("Trip to %s. Estimated Cost: $%d.", destination, proposed_cost)

  list(
    status = "SUCCESS",
    output = list(
      proposed_itinerary = itinerary,
      proposed_cost = proposed_cost,
      planner_attempts = attempts
    )
  )
})

dag$add_node(planner_node)
```

### 2. The Itinerary Validator Node

This node ensures the LLM's proposal strictly adheres to the numerical budget limit. 

```{r}
validator_node <- AgentLogicNode$new(id = "Validator", logic_fn = function(state, memory = NULL) {
  proposed_cost <- state$get("proposed_cost")
  budget <- state$get("budget")

  if (proposed_cost <= budget) {
    list(
      status = "SUCCESS",
      output = list(
        is_valid = TRUE,
        validation_message = "Itinerary meets budget requirements. Booking confirmed."
      )
    )
  } else {
    list(
      status = "SUCCESS",
      output = list(
        is_valid = FALSE,
        validation_message = sprintf("Proposed cost ($%d) exceeds budget ($%d). Find cheaper options.", proposed_cost, budget)
      )
    )
  }
})

dag$add_node(validator_node)
```

## Defining Transitions

We define the looping logic. The Validator sends control back to the Planner if `is_valid` is `FALSE`.

```{r}
dag$set_start_node("Planner")

dag$add_edge("Planner", "Validator")

dag$add_conditional_edge(
  from = "Validator",
  test = function(out) {
    isTRUE(out$is_valid)
  },
  if_true = NULL, # Ends execution successfully
  if_false = "Planner" # Loop back for a cheaper proposal
)

compiled_dag <- dag$compile()
```

## Running the Scenario

Let's execute the DAG with a strict budget of `$1500` for a trip to `Tokyo`.

```{r}
initial_state <- list(
  destination = "Tokyo",
  budget = 1500
)

cat("Starting Travel Booking Engine...\n")
result <- compiled_dag$run(initial_state = initial_state, max_steps = 10)

# Observe the outcome
cat("\n--- BOOKING RESULT ---\n")
cat("Total Planning Attempts:", result$state$get("planner_attempts"), "\n")
cat("Final Itinerary:", result$state$get("proposed_itinerary"), "\n")
cat("Validator Status:", result$state$get("validation_message"), "\n")
```

The DAG successfully cycled back through the Planner node multiple times until it generated an itinerary that satisfied the budget constraint!

<!-- APAF Bioinformatics | travel_booking.Rmd | Approved | 2026-03-29 -->
